### 2017-12-16 CWP/papers/roadmap/latex/authors.tex
Document as submitted to arXiv 2017-12-15

Alves Jr, Antonio Augusto $^{74}$;
Amadio, Guilherme $^{5}$;
Anh-Ky, Nguyen $^{111}$;
Aphecetche, Laurent $^{60}$;
Apostolakis, John $^{5}$;
Asai, Makoto $^{58,p}$;
Atzori, Luca $^{5}$;
Babik, Marian $^{5}$;
Bagliesi, Giuseppe $^{29}$;
Bandieramonte, Marilena $^{5}$;
Barisits, Martin $^{5}$;
Bauerdick, Lothar A. T. $^{16,c}$;
Belforte, Stefano $^{32}$;
Benjamin, Douglas $^{75}$;
Bernius, Catrin $^{58}$;
Bhimji, Wahid $^{42}$;
Bianchi, Riccardo Maria $^{97}$;
Bird, Ian $^{5}$;
Biscarat, Catherine $^{48}$;
Blomer, Jakob $^{5}$;
Bloom, Kenneth $^{89}$;
Boccali, Tommaso $^{29}$;
Bockelman, Brian $^{89}$;
Bold, Tomasz $^{39}$;
Bonacorsi, Daniele $^{24}$;
Boveia, Antonio $^{94}$;
Bozzi, Concezio $^{26}$;
Bracko, Marko $^{86,37}$;
Britton, David $^{79}$;
Buckley, Andy $^{79}$;
Buncic, Predrag $^{5,a}$;
Calafiura, Paolo $^{42}$;
Campana, Simone $^{5,a}$;
Canal, Philippe $^{16,c}$;
Canali, Luca $^{5}$;
Carlino, Gianpaolo $^{28}$;
Castro, Nuno $^{43,88,d}$;
Cattaneo, Marco $^{5}$;
Cerminara, Gianluca $^{5}$;
Chang, Philip $^{69}$;
Chapman, John $^{70}$;
Childers, Taylor $^{1}$;
Clarke, Peter $^{76}$;
Clemencic, Marco $^{5}$;
Cogneras, Eric $^{46}$;
Collier, Ian $^{56}$;
Corti, Gloria $^{5}$;
Cosmo, Gabriele $^{5}$;
Costanzo, Davide $^{103}$;
Couturier, Ben $^{5}$;
Cranmer, Kyle $^{53}$;
Cranshaw, Jack $^{1}$;
Crépé-Renaudin, Sabine $^{48}$;
Cristella, Leonardo $^{25}$;
Crooks, David $^{79}$;
Dallmeier-Tiessen, Sünje $^{5}$;
De Cian, Michel $^{80}$;
De, Kaushik $^{105}$;
Di Girolamo, Alessandro $^{5}$;
Dimitrov, Gancho $^{5}$;
Doglioni, Caterina $^{84,h}$;
Dotti, Andrea $^{58,p}$;
Duellmann, Dirk $^{5}$;
Duflot, Laurent $^{41}$;
Dykstra, Dave $^{16,c}$;
Dziedziniewicz-Wojcik, Katarzyna $^{5}$;
Dziurda, Agnieszka $^{5}$;
Egede, Ulrik $^{34}$;
Elmer, Peter $^{98,a}$;
Elmsheuser, Johannes $^{2}$;
Elvira, V. Daniel $^{16,c}$;
Eulisse, Giulio $^{5}$;
Ferber, Torben $^{67}$;
Filipcic, Andrej $^{37}$;
Fisk, Ian $^{59}$;
Fitzpatrick, Conor $^{14}$;
Flix, José $^{55,7,g}$;
Formica, Andrea $^{35}$;
Forti, Alessandra $^{85}$;
Gaede, Frank $^{13}$;
Ganis, Gerardo $^{5}$;
Gardner, Robert $^{73}$;
Garonne, Vincent $^{95}$;
Gellrich, Andreas $^{13}$;
Genser, Krzysztof $^{16,c}$;
George, Simon $^{57}$;
Geurts, Frank $^{99}$;
Gheata, Andrei $^{5}$;
Gheata, Mihaela $^{5}$;
Giacomini, Francesco $^{9}$;
Giagu, Stefano $^{101,31}$;
Giffels, Manuel $^{38}$;
Gingrich, Douglas $^{63}$;
Girone, Maria $^{5}$;
Gligorov, Vladimir V. $^{47}$;
Glushkov, Ivan $^{105}$;
Gohn, Wesley $^{81}$;
González Caballero, Isidro $^{96}$;
González Fernández, Juan R. $^{96}$;
Gonzalez Lopez, Jose Benito $^{5}$;
Govi, Giacomo $^{16}$;
Grandi, Claudio $^{24}$;
Grasland, Hadrien $^{41}$;
Gray, Heather $^{42}$;
Grillo, Lucia $^{85}$;
Guan, Wen $^{109}$;
Gutsche, Oliver $^{16,c}$;
Gyurjyan, Vardan $^{36}$;
Hanushevsky, Andrew $^{58,p}$;
Hariri, Farah $^{5}$;
Hartmann, Thomas $^{13}$;
Harvey, John $^{5,a}$;
Hauth, Thomas $^{38}$;
Hegner, Benedikt $^{5,a}$;
Heinemann, Beate $^{13}$;
Heinrich, Lukas $^{53}$;
Hernández, José M. $^{7,g}$;
Hildreth, Michael $^{91,f}$;
Hodgkinson, Mark $^{103}$;
Hoeche, Stefan $^{58,p}$;
Hristov, Peter $^{5}$;
Huang, Xingtao $^{102}$;
Ivanchenko, Vladimir N. $^{5,106}$;
Ivanov, Todor $^{104}$;
Jashal, Brij $^{62}$;
Jayatilaka, Bodhitha $^{16,c}$;
Jones, Roger $^{82,a}$;
Jouvin, Michel $^{41,a}$;
Jun, Soon Yung $^{16,c}$;
Kagan, Michael $^{58,p}$;
Kalderon, Charles William $^{84}$;
Karavakis, Edward $^{5}$;
Katz, Daniel S. $^{72}$;
Kcira, Dorian $^{11}$;
Kersevan, Borut Paul $^{83}$;
Kirby, Michael $^{16,c}$;
Klimentov, Alexei $^{2}$;
Klute, Markus $^{49}$;
Komarov, Ilya $^{32,n}$;
Koppenburg, Patrick $^{52}$;
Kowalkowski, Jim $^{16,c}$;
Kreczko, Luke $^{66}$;
Kuhr, Thomas $^{45,a}$;
Kutschke, Robert $^{16,a,c}$;
Kuznetsov, Valentin $^{12}$;
Lampl, Walter $^{65}$;
Lancon, Eric $^{2}$;
Lange, David $^{98}$;
Lassnig, Mario $^{5}$;
Laycock, Paul $^{5}$;
Leggett, Charles $^{42}$;
Letts, James $^{69}$;
Lewendel, Birgit $^{13}$;
Li, Teng $^{76}$;
Lima, Guilherme $^{16}$;
Linacre, Jacob $^{56,m}$;
Linden, Tomas $^{18}$;
Lo Presti, Giuseppe $^{5}$;
Lopienski, Sebastian $^{5}$;
Love, Peter $^{82}$;
Marshall, Zachary L $^{42}$;
Martelli, Edoardo $^{5}$;
Martin-Haugh, Stewart $^{56}$;
Mato, Pere $^{5}$;
Mazumdar, Kajari $^{62}$;
McCauley, Thomas $^{91}$;
McFayden, Josh $^{5}$;
McKee, Shawn $^{87,l}$;
McNab, Andrew $^{85}$;
Meinhard, Helge $^{5}$;
Menasce, Dario $^{27,a}$;
Mendez Lorenzo, Patricia $^{5}$;
Mete, Alaettin Serhan $^{68}$;
Michelotto, Michele $^{30}$;
Mitrevski, Jovan $^{45}$;
Moneta, Lorenzo $^{5}$;
Morgan, Ben $^{108}$;
Mount, Richard $^{58,p}$;
Moyse, Edward $^{64}$;
Murray, Sean $^{71,10}$;
Neubauer, Mark S $^{72,a,k}$;
Novaes, Sérgio $^{100}$;
Novak, Mihaly $^{5}$;
Oyanguren, Arantza $^{21}$;
Ozturk, Nurcan $^{105}$;
Pacheco Pages, Andres $^{55,19,j}$;
Paganini, Michela $^{110}$;
Pansanel, Jerome $^{33}$;
Pascuzzi, Vincent R. $^{107}$;
Pearce, Alex $^{5}$;
Pearson, Ben $^{50}$;
Pedro, Kevin $^{16,c}$;
Perdue, Gabriel $^{16}$;
Perez-Calero Yzquierdo, Antonio $^{55,7,g}$;
Perrozzi, Luca $^{15}$;
Petersen, Troels $^{51}$;
Petric, Marko $^{5}$;
Piedra, Jónatan $^{20}$;
Piilonen, Leo $^{112,i}$;
Piparo, Danilo $^{5}$;
Pokorski, Witold $^{5}$;
Polci, Francesco $^{47}$;
Potamianos, Karolos $^{13}$;
Psihas, Fernanda $^{23}$;
Raven, Gerhard $^{52}$;
Reuter, Jürgen $^{13}$;
Ribon, Alberto $^{5}$;
Ritter, Martin $^{45}$;
Robinson, James $^{13}$;
Rodrigues, Eduardo $^{74,a,e}$;
Roiser, Stefan $^{5,a}$;
Rousseau, David $^{41}$;
Roy, Gareth $^{79}$;
Sailer, Andre $^{5}$;
Sakuma, Tai $^{66}$;
Sánchez-Hernández, Alberto $^{8,o}$;
Santana, Renato $^{3}$;
Sartirana, Andrea $^{44}$;
Schellman, Heidi $^{54}$;
Schovancová, Jaroslava $^{5}$;
Schramm, Steven $^{78}$;
Schulz, Markus $^{5}$;
Sciabà, Andrea $^{5}$;
Seidel, Sally $^{90}$;
Sekmen, Sezen $^{40}$;
Serfon, Cedric $^{93}$;
Severini, Horst $^{92}$;
Sexton-Kennedy, Elizabeth $^{16,a,c}$;
Seymour, Michael $^{85}$;
Shapoval, Illya $^{42}$;
Shiers, Jamie $^{5}$;
Shiu, Jing-Ge $^{61}$;
Short, Hannah $^{5}$;
Siroli, Gian Piero $^{24}$;
Skipsey, Sam $^{79}$;
Smith, Tim $^{5}$;
Snyder, Scott $^{2}$;
Sokoloff, Michael D $^{74,a}$;
Stadie, Hartmut $^{17}$;
Stark, Giordon $^{6}$;
Stewart, Gordon $^{79}$;
Stewart, Graeme $^{5,a}$;
Templon, Jeff $^{52}$;
Tenaglia, Giacomo $^{5}$;
Tsulaia, Vakhtang $^{42}$;
Tunnell, Christopher $^{6}$;
Vaandering, Eric $^{16,c}$;
Valassi, Andrea $^{5}$;
Vallecorsa, Sofia $^{77}$;
Valsan, Liviu $^{5}$;
Van Gemmeren, Peter $^{1}$;
Vázquez Sierra, Carlos $^{52}$;
Vernet, Renaud $^{4}$;
Viren, Brett $^{2}$;
Vlimant, Jean-Roch $^{11,a}$;
Voss, Christian $^{13}$;
Vuosalo, Carl $^{113}$;
Wartel, Romain $^{5}$;
Wenaus, Torre $^{2}$;
Wenzel, Sandro $^{5}$;
Winklmeier, Frank $^{54}$;
Wissing, Christoph $^{13}$;
Wuerthwein, Frank $^{69}$;
Wynne, Benjamin $^{76}$;
Xiaomei, Zhang $^{22}$;
Yang, Wei $^{58,p}$;
Yazgan, Efe $^{22}$
\bigskip
\par {\footnotesize $^{1}$ High Energy Physics Division, Argonne National Laboratory, Argonne IL, USA}
\par {\footnotesize $^{2}$ Physics Department, Brookhaven National Laboratory, Upton NY, USA}
\par {\footnotesize $^{3}$ Centro Brasileiro de Pesquisas Fisicas, Rio de Janeiro, Brazil}
\par {\footnotesize $^{4}$ Centre de Calcul de l’IN2P3, Villeurbanne, Lyon, France}
\par {\footnotesize $^{5}$ CERN, Geneva, Switzerland}
\par {\footnotesize $^{6}$ Enrico Fermi Institute, University of Chicago, Chicago IL, USA}
\par {\footnotesize $^{7}$ Centro de Investigaciones Energéticas Medioambientales y Tecnológicas (CIEMAT), Madrid, Spain}
\par {\footnotesize $^{8}$ Cinvestav, Mexico City, Mexico}
\par {\footnotesize $^{9}$ Centro Nazionale Analisi Fotogrammi (CNAF), INFN, Bologna, Italy}
\par {\footnotesize $^{10}$ Center for High Performance Computing, Cape Town, South Africa}
\par {\footnotesize $^{11}$ California Institute of Technology, Pasadena, California, USA}
\par {\footnotesize $^{12}$ Cornell University, Ithaca, USA}
\par {\footnotesize $^{13}$ Deutsches Elektronen-Synchrotron, Hamburg, Germany}
\par {\footnotesize $^{14}$ Institute of Physics, Ecole Polytechnique Federale de Lausanne (EPFL), Lausanne, Switzerland}
\par {\footnotesize $^{15}$ ETH Zurich - Institute for Particle Physics and Astrophysics (IPA), Zurich, Switzerland}
\par {\footnotesize $^{16}$ Fermi National Accelerator Laboratory, Batavia, USA}
\par {\footnotesize $^{17}$ University of Hamburg, Hamburg, Germany}
\par {\footnotesize $^{18}$ Helsinki Institute of Physics, Helsinki, Finland}
\par {\footnotesize $^{19}$ Institut de Física d’Altes Energies and Departament de Física de la Universitat Autònoma de Barcelona and ICREA, Barcelona, Spain}
\par {\footnotesize $^{20}$ Instituto de Física de Cantabria (IFCA), CSIC-Universidad de Cantabria, Santander, Spain}
\par {\footnotesize $^{21}$ Instituto de Fisica Corpuscular, Centro Mixto Universidad de Valencia - CSIC, Valencia, Spain}
\par {\footnotesize $^{22}$ Institute of High Energy Physics, Chinese Academy of Sciences, Beijing}
\par {\footnotesize $^{23}$ Department of Physics, Indiana University, Bloomington IN, USA}
\par {\footnotesize $^{24}$ INFN Sezione di Bologna, Università di Bologna, Bologna, Italy}
\par {\footnotesize $^{25}$ INFN Sezione di Bari, Università di Bari, Politecnico di Bari, Bari, Italy}
\par {\footnotesize $^{26}$ Universita e INFN, Ferrara, Ferrara, Italy}
\par {\footnotesize $^{27}$ INFN Sezione di Milano-Bicocca, Milano, Italy}
\par {\footnotesize $^{28}$ INFN Sezione di Napoli, Università di Napoli, Napoli, Italy}
\par {\footnotesize $^{29}$ INFN Sezione di Pisa, Università di Pisa, Scuola Normale Superiore di Pisa, Pisa, Italy}
\par {\footnotesize $^{30}$ INFN Sezione di Padova, Università di Padova b, Padova, Italy}
\par {\footnotesize $^{31}$ INFN Sezione di Roma I, Università La Sapienza, Roma, Italy}
\par {\footnotesize $^{32}$ INFN Sezione di Trieste, Università di Trieste, Trieste, Italy}
\par {\footnotesize $^{33}$ Université de Strasbourg, CNRS, IPHC UMR 7178, F-67000 Strasbourg, France}
\par {\footnotesize $^{34}$ Imperial College London, London, United Kingdom}
\par {\footnotesize $^{35}$ DSM/IRFU (Institut de Recherches sur les Lois Fondamentales de l’Univers), CEA Saclay (Commissariat a l’Energie Atomique), Gif-sur-Yvette, France}
\par {\footnotesize $^{36}$ Thomas Jefferson National Accelerator Facility, Newport News, Virginia, USA}
\par {\footnotesize $^{37}$ Jozef Stefan Institute, Ljubljana, Slovenia}
\par {\footnotesize $^{38}$ Karlsruhe Institute of Technology, Karlsruhe, Germany}
\par {\footnotesize $^{39}$ AGH University of Science and Technology, Faculty of Physics and Applied Computer Science, Krakow, Poland}
\par {\footnotesize $^{40}$ Kyungpook National University, Daegu, Republic of Korea}
\par {\footnotesize $^{41}$ LAL, Université Paris-Sud and CNRS/IN2P3, Orsay, France}
\par {\footnotesize $^{42}$ Lawrence Berkeley National Laboratory and University of California, Berkeley CA, USA}
\par {\footnotesize $^{43}$ Laboratorio de Instrumentacao e Fisica Experimental de Particulas (LIP), Lisboa, Portugal}
\par {\footnotesize $^{44}$ Laboratoire Leprince-Ringuet, Ecole polytechnique, CNRS/IN2P3, Université Paris-Saclay, Palaiseau, France}
\par {\footnotesize $^{45}$ Fakultät für Physik, Ludwig-Maximilians-Universität München, München, Germany}
\par {\footnotesize $^{46}$ Laboratoire de Physique Corpusculaire, Clermont Université and Université Blaise Pascal and CNRS/IN2P3, Clermont-Ferrand, France}
\par {\footnotesize $^{47}$ LPNHE, Universite Pierre et Marie Curie, Universite Paris Diderot, CNRS/IN2P3, Paris, France}
\par {\footnotesize $^{48}$ Laboratoire de Physique Subatomique et de Cosmologie, Université Joseph Fourier and CNRS/IN2P3 and Institut National Polytechnique de Grenoble, Grenoble, France}
\par {\footnotesize $^{49}$ Department of Physics, University of Massachusetts, Amherst, MA, USA}
\par {\footnotesize $^{50}$ Max-Planck-Institut für Physik (Werner-Heisenberg-Institut), München, Germany}
\par {\footnotesize $^{51}$ Niels Bohr Institute, University of Copenhagen, Kobenhavn, Denmark}
\par {\footnotesize $^{52}$ Nikhef National Institute for Subatomic Physics and University of Amsterdam, Amsterdam, Netherlands}
\par {\footnotesize $^{53}$ Department of Physics, New York University, New York, NY, USA}
\par {\footnotesize $^{54}$ Center for High Energy Physics, University of Oregon, Eugene OR, USA}
\par {\footnotesize $^{55}$ Port d’Informació Científica (PIC), Universitat Autònoma de Barcelona (UAB), Barcelona, Spain}
\par {\footnotesize $^{56}$ STFC Rutherford Appleton Laboratory, Didcot, United Kingdom}
\par {\footnotesize $^{57}$ Department of Physics, Royal Holloway University of London, Surrey, United Kingdom}
\par {\footnotesize $^{58}$ SLAC National Accelerator Laboratory, Menlo Park CA, USA}
\par {\footnotesize $^{59}$ Simons Foundation, New York, USA}
\par {\footnotesize $^{60}$ SUBATECH, IMT Atlantique, Université de Nantes, CNRS-IN2P3, Nantes, France}
\par {\footnotesize $^{61}$ National Taiwan University, Taipei, Taiwan}
\par {\footnotesize $^{62}$ Tata Institute of Fundamental Research, Mumbai, India}
\par {\footnotesize $^{63}$ Department of Physics, University of Alberta, Edmonton, AB, Canada}
\par {\footnotesize $^{64}$ Department of Physics, University of Massachusetts, Amherst, MA, USA}
\par {\footnotesize $^{65}$ Department of Physics, University of Arizona, Tucson, AZ, USA}
\par {\footnotesize $^{66}$ H.H. Wills Physics Laboratory, University of Bristol, Bristol, United Kingdom}
\par {\footnotesize $^{67}$ Department of Physics, University of British Columbia, Vancouver, BC, Canada}
\par {\footnotesize $^{68}$ Department of Physics and Astronomy, University of California Irvine, Irvine, CA, USA}
\par {\footnotesize $^{69}$ University of California, San Diego, La Jolla, USA}
\par {\footnotesize $^{70}$ Cavendish Laboratory, University of Cambridge, Cambridge, United Kingdom}
\par {\footnotesize $^{71}$ Physics Department, University of Cape Town, Cape Town, South Africa}
\par {\footnotesize $^{72}$ University of Illinois Urbana-Champaign, Champaign, Illinois, USA}
\par {\footnotesize $^{73}$ Enrico Fermi Institute, University of Chicago, Chicago, IL, USA}
\par {\footnotesize $^{74}$ University of Cincinnati, Cincinnati, OH, USA}
\par {\footnotesize $^{75}$ Department of Physics, Duke University, Durham, NC, USA}
\par {\footnotesize $^{76}$ SUPA - School of Physics and Astronomy, University of Edinburgh, Edinburgh, United Kingdom}
\par {\footnotesize $^{77}$ Gangneung-Wonju National University, South Korea}
\par {\footnotesize $^{78}$ Section de Physique, Université de Genève, Geneva, Switzerland}
\par {\footnotesize $^{79}$ SUPA - School of Physics and Astronomy, University of Glasgow, Glasgow, United Kingdom}
\par {\footnotesize $^{80}$ Physikalisches Institut, Ruprecht-Karls-Universitat Heidelberg, Heidelberg, Germany}
\par {\footnotesize $^{81}$ Department of Physics \& Astronomy, University of Kentucky, Lexington, USA}
\par {\footnotesize $^{82}$ Physics Department, Lancaster University, Lancaster, United Kingdom}
\par {\footnotesize $^{83}$ Department of Physics, Jožef Stefan Institute and University of Ljubljana, Ljubljana, Slovenia}
\par {\footnotesize $^{84}$ Fysiska institutionen, Lunds universitet, Lund, Sweden}
\par {\footnotesize $^{85}$ School of Physics and Astronomy, University of Manchester, Manchester, United Kingdom}
\par {\footnotesize $^{86}$ University of Maribor, Ljubljana, Slovenia}
\par {\footnotesize $^{87}$ Department of Physics, The University of Michigan, Ann Arbor, MI, USA}
\par {\footnotesize $^{88}$ Departamento de Fisica, Universidade de Minho, Braga, Portugal}
\par {\footnotesize $^{89}$ University of Nebraska-Lincoln, Lincoln, USA}
\par {\footnotesize $^{90}$ Department of Physics and Astronomy, University of New Mexico, Albuquerque NM, USA}
\par {\footnotesize $^{91}$ University of Notre Dame, Notre Dame, USA}
\par {\footnotesize $^{92}$ Homer L. Dodge Department of Physics and Astronomy, University of Oklahoma, Norman OK, USA}
\par {\footnotesize $^{93}$ Department of Physics, University of Oslo, Oslo, Norway}
\par {\footnotesize $^{94}$ The Ohio State University, Columbus, USA}
\par {\footnotesize $^{95}$ Department of Physics, University of Oslo, Oslo, Norway}
\par {\footnotesize $^{96}$ Universidad de Oviedo, Oviedo, Spain}
\par {\footnotesize $^{97}$ Department of Physics and Astronomy, University of Pittsburgh, Pittsburgh, PA, USA}
\par {\footnotesize $^{98}$ Princeton University, Princeton, USA}
\par {\footnotesize $^{99}$ Rice University, Houston, USA}
\par {\footnotesize $^{100}$ Universidade Estadual Paulista, São Paulo, Brazil}
\par {\footnotesize $^{101}$ Dipartimento di Fisica, Università La Sapienza, Roma, Italy}
\par {\footnotesize $^{102}$ School of Physics, Shandong University, Shandong, China}
\par {\footnotesize $^{103}$ Department of Physics and Astronomy, University of Sheffield, Sheffield, United Kingdom}
\par {\footnotesize $^{104}$ University of Sofia, Sofia, Bulgaria}
\par {\footnotesize $^{105}$ Department of Physics, The University of Texas at Arlington, Arlington, TX, USA}
\par {\footnotesize $^{106}$ National Research Tomsk Polytechnic University, Tomsk, Russia}
\par {\footnotesize $^{107}$ Department of Physics, University of Toronto, Toronto, ON, Canada}
\par {\footnotesize $^{108}$ Department of Physics, University of Warwick, Coventry, United Kingdom}
\par {\footnotesize $^{109}$ University ofWisconsin - Madison, Madison, WI, USA}
\par {\footnotesize $^{110}$ Department of Physics, Yale University, New Haven, CT, USA}
\par {\footnotesize $^{111}$ IOP and GUST, Vietnam Academy of Science and Technology (VAST), Hanoi, Vietnam}
\par {\footnotesize $^{112}$ Virginia Tech, Blacksburg, Virginia, USA}
\par {\footnotesize $^{113}$ Department of Physics, University of Wisconsin, Madison WI, USA}
\bigskip
\par {\footnotesize $^{a}$ Community White Paper Editorial Board Member}
\par {\footnotesize $^{b}$ Vladimir V. Gligorov acknowledges funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme under grant agreement No 724777 “RECEPT”.}
\par {\footnotesize $^{c}$ Supported by the US-DOE, DE-AC02-07CH11359}
\par {\footnotesize $^{d}$ Supported by FCT-Portugal, IF/00050/2013/CP1172/CT0002}
\par {\footnotesize $^{e}$ Supported by the US-NSF, ACI-1450319}
\par {\footnotesize $^{f}$ Supported by the US-NSF, PHY-1607578}
\par {\footnotesize $^{g}$ Supported by ES-MINECO, FPA2016-80994-c2-1-R \& MDM-2015-0509}
\par {\footnotesize $^{h}$ Caterina Doglioni acknowledges funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme under grant agreement No 679305 “DARKJETS”.}
\par {\footnotesize $^{i}$ Supported by the US-DOE, DE-SC0009973}
\par {\footnotesize $^{j}$ Supported by the ES-MINECO FPA2016-80994-C2-2-R}
\par {\footnotesize $^{k}$ Supported by the US-DOE, DE-SC0018098 and US-NSF ACI-1558233}
\par {\footnotesize $^{l}$ Supported by the US-DOE, DE-SC0007859 and US-NSF, 76749/1136652/2}
\par {\footnotesize $^{m}$ Supported by funding from the European Union’s Horizon 2020 research and innovation programme under the Marie Skłodowska-Curie grant agreement No. 752730}
\par {\footnotesize $^{n}$ Supported by Swiss National Science Foundation, Early Postdoc Mobility Fellowship, project number P2ELP2\_168556}
\par {\footnotesize $^{o}$ Supported by CONACYT (Mexico)}
\par {\footnotesize $^{p}$ Supported by the US-DOE, DE-AC02-76SF0051}