### 2018-01-16 CWP/papers/HSF-CWP-2017-01_roadmap/latex/authors.tex

Albrecht, Johannes$^{66}$;
Alves Jr, Antonio Augusto$^{79}$;
Amadio, Guilherme$^{5}$;
Anh-Ky, Nguyen$^{117}$;
Aphecetche, Laurent$^{63}$;
Apostolakis, John$^{5}$;
Asai, Makoto$^{61,p}$;
Atzori, Luca$^{5}$;
Babik, Marian$^{5}$;
Bagliesi, Giuseppe$^{31}$;
Bandieramonte, Marilena$^{5}$;
Barisits, Martin$^{5}$;
Bauerdick, Lothar A. T.$^{16,c}$;
Belforte, Stefano$^{34}$;
Benjamin, Douglas$^{80}$;
Bernius, Catrin$^{61}$;
Bhimji, Wahid$^{44}$;
Bianchi, Riccardo Maria$^{102}$;
Bird, Ian$^{5}$;
Biscarat, Catherine$^{50}$;
Blomer, Jakob$^{5}$;
Bloom, Kenneth$^{94}$;
Boccali, Tommaso$^{31}$;
Bockelman, Brian$^{94}$;
Bold, Tomasz$^{41}$;
Bonacorsi, Daniele$^{25}$;
Boveia, Antonio$^{98}$;
Bozzi, Concezio$^{27}$;
Bracko, Marko$^{91,39}$;
Britton, David$^{84}$;
Buckley, Andy$^{84}$;
Buncic, Predrag$^{5,a}$;
Calafiura, Paolo$^{44}$;
Campana, Simone$^{5,a}$;
Canal, Philippe$^{16,c}$;
Canali, Luca$^{5}$;
Carlino, Gianpaolo$^{30}$;
Castro, Nuno$^{45,93,d}$;
Cattaneo, Marco$^{5}$;
Cerminara, Gianluca$^{5}$;
Chang, Philip$^{73}$;
Chapman, John$^{74}$;
Chen, Gang$^{23}$;
Childers, Taylor$^{1}$;
Clarke, Peter$^{81}$;
Clemencic, Marco$^{5}$;
Cogneras, Eric$^{48}$;
Collier, Ian$^{59}$;
Corti, Gloria$^{5}$;
Cosmo, Gabriele$^{5}$;
Costanzo, Davide$^{108}$;
Couturier, Ben$^{5}$;
Cranmer, Kyle$^{55}$;
Cranshaw, Jack$^{1}$;
Cristella, Leonardo$^{26}$;
Crooks, David$^{84}$;
Crépé-Renaudin, Sabine$^{50}$;
Dallmeier-Tiessen, Sünje$^{5}$;
De, Kaushik$^{110}$;
De Cian, Michel$^{85}$;
Di Girolamo, Alessandro$^{5}$;
Dimitrov, Gancho$^{5}$;
Doglioni, Caterina$^{89,h}$;
Dotti, Andrea$^{61,p}$;
Duellmann, Dirk$^{5}$;
Duflot, Laurent$^{43}$;
Dykstra, Dave$^{16,c}$;
Dziedziniewicz-Wojcik, Katarzyna$^{5}$;
Dziurda, Agnieszka$^{5}$;
Egede, Ulrik$^{36}$;
Elmer, Peter$^{103,a}$;
Elmsheuser, Johannes$^{2}$;
Elvira, V. Daniel$^{16,c}$;
Eulisse, Giulio$^{5}$;
Ferber, Torben$^{71}$;
Filipcic, Andrej$^{39}$;
Fisk, Ian$^{62}$;
Fitzpatrick, Conor$^{14}$;
Flix, José$^{57,7,g}$;
Formica, Andrea$^{37}$;
Forti, Alessandra$^{90}$;
Frost, James$^{101}$;
Gaede, Frank$^{13}$;
Ganis, Gerardo$^{5}$;
Gardner, Robert$^{78}$;
Garonne, Vincent$^{99}$;
Gellrich, Andreas$^{13}$;
Genser, Krzysztof$^{16,c}$;
George, Simon$^{60}$;
Geurts, Frank$^{104}$;
Gheata, Andrei$^{5}$;
Gheata, Mihaela$^{5}$;
Giacomini, Francesco$^{9}$;
Giagu, Stefano$^{106,33}$;
Giffels, Manuel$^{40}$;
Gingrich, Douglas$^{67}$;
Girone, Maria$^{5}$;
Gligorov, Vladimir V.$^{49}$;
Glushkov, Ivan$^{110}$;
Gohn, Wesley$^{86}$;
Gonzalez Lopez, Jose Benito$^{5}$;
González Caballero, Isidro$^{100}$;
González Fernández, Juan R.$^{100}$;
Govi, Giacomo$^{16}$;
Grandi, Claudio$^{25}$;
Grasland, Hadrien$^{43}$;
Gray, Heather$^{44}$;
Grillo, Lucia$^{90}$;
Guan, Wen$^{115}$;
Gutsche, Oliver$^{16,c}$;
Gyurjyan, Vardan$^{38}$;
Hanushevsky, Andrew$^{61,p}$;
Hariri, Farah$^{5}$;
Hartmann, Thomas$^{13}$;
Harvey, John$^{5,a}$;
Hauth, Thomas$^{40}$;
Hegner, Benedikt$^{5,a}$;
Heinemann, Beate$^{13}$;
Heinrich, Lukas$^{55}$;
Hernández, José M.$^{7,g}$;
Hildreth, Michael$^{96,f}$;
Hodgkinson, Mark$^{108}$;
Hoeche, Stefan$^{61,p}$;
Hristov, Peter$^{5}$;
Huang, Xingtao$^{107}$;
Ivanchenko, Vladimir N.$^{5,111}$;
Ivanov, Todor$^{109}$;
Jashal, Brij$^{65}$;
Jayatilaka, Bodhitha$^{16,c}$;
Jones, Roger$^{87,a}$;
Jouvin, Michel$^{43,a}$;
Jun, Soon Yung$^{16,c}$;
Kagan, Michael$^{61,p}$;
Kalderon, Charles William$^{89}$;
Karavakis, Edward$^{5}$;
Katz, Daniel S.$^{77}$;
Kcira, Dorian$^{11}$;
Kersevan, Borut Paul$^{88}$;
Kirby, Michael$^{16,c}$;
Klimentov, Alexei$^{2}$;
Klute, Markus$^{51}$;
Komarov, Ilya$^{34,n}$;
Konstantinov, Dmitri$^{58}$;
Koppenburg, Patrick$^{54}$;
Kowalkowski, Jim$^{16,c}$;
Kreczko, Luke$^{70}$;
Kuhr, Thomas$^{47,a}$;
Kutschke, Robert$^{16,a,c}$;
Kuznetsov, Valentin$^{12}$;
Lampl, Walter$^{69}$;
Lancon, Eric$^{2}$;
Lange, David$^{103,a}$;
Lassnig, Mario$^{5}$;
Laycock, Paul$^{5}$;
Leggett, Charles$^{44}$;
Letts, James$^{73}$;
Lewendel, Birgit$^{13}$;
Li, Teng$^{81}$;
Lima, Guilherme$^{16}$;
Linacre, Jacob$^{59,m}$;
Linden, Tomas$^{18}$;
Lo Presti, Giuseppe$^{5}$;
Lopienski, Sebastian$^{5}$;
Love, Peter$^{87}$;
Magini, Nicolò$^{28}$;
Marshall, Zachary L$^{44}$;
Martelli, Edoardo$^{5}$;
Martin-Haugh, Stewart$^{59}$;
Mato, Pere$^{5}$;
Mazumdar, Kajari$^{65}$;
McCauley, Thomas$^{96}$;
McFayden, Josh$^{5}$;
McKee, Shawn$^{92,l}$;
McNab, Andrew$^{90}$;
Mehdiyev, Rashid$^{76}$;
Meinhard, Helge$^{5}$;
Menasce, Dario$^{29,a}$;
Mendez Lorenzo, Patricia$^{5}$;
Mete, Alaettin Serhan$^{72}$;
Michelotto, Michele$^{32}$;
Mitrevski, Jovan$^{47}$;
Moneta, Lorenzo$^{5}$;
Morgan, Ben$^{114}$;
Mount, Richard$^{61,p}$;
Moyse, Edward$^{68}$;
Murray, Sean$^{75,10}$;
Neubauer, Mark S$^{77,a,k}$;
Novaes, Sérgio$^{105}$;
Novak, Mihaly$^{5}$;
Oyanguren, Arantza$^{22}$;
Ozturk, Nurcan$^{110}$;
Pacheco Pages, Andres$^{57,20,j}$;
Paganini, Michela$^{116}$;
Pansanel, Jerome$^{35}$;
Pascuzzi, Vincent R.$^{112}$;
Patrick, Glenn$^{59}$;
Pearce, Alex$^{5}$;
Pearson, Ben$^{52}$;
Pedro, Kevin$^{16,c}$;
Perdue, Gabriel$^{16}$;
Perez-Calero Yzquierdo, Antonio$^{57,7,g}$;
Perrozzi, Luca$^{15}$;
Petersen, Troels$^{53}$;
Petric, Marko$^{5}$;
Piedra, Jónatan$^{21}$;
Piilonen, Leo$^{118,i}$;
Piparo, Danilo$^{5}$;
Pokorski, Witold$^{5}$;
Polci, Francesco$^{49}$;
Potamianos, Karolos$^{13}$;
Psihas, Fernanda$^{24}$;
Raven, Gerhard$^{54}$;
Reuter, Jürgen$^{13}$;
Ribon, Alberto$^{5}$;
Rinaldi, Lorenzo$^{25}$;
Ritter, Martin$^{47}$;
Robinson, James$^{13}$;
Rodrigues, Eduardo$^{79,a,e}$;
Roiser, Stefan$^{5,a}$;
Rousseau, David$^{43}$;
Roy, Gareth$^{84}$;
Sailer, Andre$^{5}$;
Sakuma, Tai$^{70}$;
Santana, Renato$^{3}$;
Sartirana, Andrea$^{46}$;
Schellman, Heidi$^{56}$;
Schovancová, Jaroslava$^{5}$;
Schramm, Steven$^{83}$;
Schulz, Markus$^{5}$;
Sciabà, Andrea$^{5}$;
Seidel, Sally$^{95}$;
Sekmen, Sezen$^{42}$;
Serfon, Cedric$^{99}$;
Severini, Horst$^{97}$;
Sexton-Kennedy, Elizabeth$^{16,a,c}$;
Seymour, Michael$^{90}$;
Shapoval, Illya$^{44}$;
Shiers, Jamie$^{5}$;
Shiu, Jing-Ge$^{64}$;
Short, Hannah$^{5}$;
Siroli, Gian Piero$^{25}$;
Skipsey, Sam$^{84}$;
Smith, Tim$^{5}$;
Snyder, Scott$^{2}$;
Sokoloff, Michael D$^{79,a}$;
Stadie, Hartmut$^{17}$;
Stark, Giordon$^{6}$;
Stewart, Gordon$^{84}$;
Stewart, Graeme$^{5,a}$;
Sánchez, Arturo$^{113,19}$;
Sánchez-Hernández, Alberto$^{8,o}$;
Taffard, Anyes$^{72}$;
Templon, Jeff$^{54}$;
Tenaglia, Giacomo$^{5}$;
Tsulaia, Vakhtang$^{44}$;
Tunnell, Christopher$^{6}$;
Vaandering, Eric$^{16,c}$;
Valassi, Andrea$^{5}$;
Vallecorsa, Sofia$^{82}$;
Valsan, Liviu$^{5}$;
Van Gemmeren, Peter$^{1}$;
Vernet, Renaud$^{4}$;
Viren, Brett$^{2}$;
Vlimant, Jean-Roch$^{11,a}$;
Voss, Christian$^{13}$;
Vuosalo, Carl$^{115}$;
Vázquez Sierra, Carlos$^{54}$;
Wartel, Romain$^{5}$;
Wenaus, Torre$^{2}$;
Wenzel, Sandro$^{5}$;
Winklmeier, Frank$^{56}$;
Wissing, Christoph$^{13}$;
Wuerthwein, Frank$^{73}$;
Wynne, Benjamin$^{81}$;
Xiaomei, Zhang$^{23}$;
Yang, Wei$^{61,p}$;
Yazgan, Efe$^{23}$
\bigskip
\par {\footnotesize $^{1}$ High Energy Physics Division, Argonne National Laboratory, Argonne IL, USA}
\par {\footnotesize $^{2}$ Physics Department, Brookhaven National Laboratory, Upton NY, USA}
\par {\footnotesize $^{3}$ Centro Brasileiro de Pesquisas Físicas, Rio de Janeiro, Brazil}
\par {\footnotesize $^{4}$ Centre de Calcul de l’IN2P3, Villeurbanne, Lyon, France}
\par {\footnotesize $^{5}$ CERN, Geneva, Switzerland}
\par {\footnotesize $^{6}$ Enrico Fermi Institute, University of Chicago, Chicago IL, USA}
\par {\footnotesize $^{7}$ Centro de Investigaciones Energéticas Medioambientales y Tecnológicas (CIEMAT), Madrid, Spain}
\par {\footnotesize $^{8}$ Cinvestav, Mexico City, Mexico}
\par {\footnotesize $^{9}$ Centro Nazionale Analisi Fotogrammi (CNAF), INFN, Bologna, Italy}
\par {\footnotesize $^{10}$ Center for High Performance Computing, Cape Town, South Africa}
\par {\footnotesize $^{11}$ California Institute of Technology, Pasadena, California, USA}
\par {\footnotesize $^{12}$ Cornell University, Ithaca, USA}
\par {\footnotesize $^{13}$ Deutsches Elektronen-Synchrotron, Hamburg, Germany}
\par {\footnotesize $^{14}$ Institute of Physics, École Polytechnique Fédérale de Lausanne (EPFL), Lausanne, Switzerland}
\par {\footnotesize $^{15}$ ETH Zurich - Institute for Particle Physics and Astrophysics (IPA), Zurich, Switzerland}
\par {\footnotesize $^{16}$ Fermi National Accelerator Laboratory, Batavia, USA}
\par {\footnotesize $^{17}$ University of Hamburg, Hamburg, Germany}
\par {\footnotesize $^{18}$ Helsinki Institute of Physics, Helsinki, Finland}
\par {\footnotesize $^{19}$ International Center for Theoretical Physics, Trieste, Italy}
\par {\footnotesize $^{20}$ Institut de Física d’Altes Energies and Departament de Física de la Universitat Autònoma de Barcelona and ICREA, Barcelona, Spain}
\par {\footnotesize $^{21}$ Instituto de Física de Cantabria (IFCA), CSIC-Universidad de Cantabria, Santander, Spain}
\par {\footnotesize $^{22}$ Instituto de Física Corpuscular, Centro Mixto Universidad de Valencia - CSIC, Valencia, Spain}
\par {\footnotesize $^{23}$ Institute of High Energy Physics, Chinese Academy of Sciences, Beijing}
\par {\footnotesize $^{24}$ Department of Physics, Indiana University, Bloomington IN, USA}
\par {\footnotesize $^{25}$ INFN Sezione di Bologna, Università di Bologna, Bologna, Italy}
\par {\footnotesize $^{26}$ INFN Sezione di Bari, Università di Bari, Politecnico di Bari, Bari, Italy}
\par {\footnotesize $^{27}$ Università e INFN, Ferrara, Ferrara, Italy}
\par {\footnotesize $^{28}$ INFN Sezione di Genova, Genova, Italy}
\par {\footnotesize $^{29}$ INFN Sezione di Milano-Bicocca, Milano, Italy}
\par {\footnotesize $^{30}$ INFN Sezione di Napoli, Università di Napoli, Napoli, Italy}
\par {\footnotesize $^{31}$ INFN Sezione di Pisa, Università di Pisa, Scuola Normale Superiore di Pisa, Pisa, Italy}
\par {\footnotesize $^{32}$ INFN Sezione di Padova, Università di Padova b, Padova, Italy}
\par {\footnotesize $^{33}$ INFN Sezione di Roma I, Università La Sapienza, Roma, Italy}
\par {\footnotesize $^{34}$ INFN Sezione di Trieste, Università di Trieste, Trieste, Italy}
\par {\footnotesize $^{35}$ Université de Strasbourg, CNRS, IPHC UMR 7178, F-67000 Strasbourg, France}
\par {\footnotesize $^{36}$ Imperial College London, London, United Kingdom}
\par {\footnotesize $^{37}$ DSM/IRFU (Institut de Recherches sur les Lois Fondamentales de l’Univers), CEA Saclay (Commissariat à l’Énergie Atomique), Gif-sur-Yvette, France}
\par {\footnotesize $^{38}$ Thomas Jefferson National Accelerator Facility, Newport News, Virginia, USA}
\par {\footnotesize $^{39}$ Jožef Stefan Institute, Ljubljana, Slovenia}
\par {\footnotesize $^{40}$ Karlsruhe Institute of Technology, Karlsruhe, Germany}
\par {\footnotesize $^{41}$ AGH University of Science and Technology, Faculty of Physics and Applied Computer Science, Krakow, Poland}
\par {\footnotesize $^{42}$ Kyungpook National University, Daegu, Republic of Korea}
\par {\footnotesize $^{43}$ LAL, Université Paris-Sud and CNRS/IN2P3, Orsay, France}
\par {\footnotesize $^{44}$ Lawrence Berkeley National Laboratory and University of California, Berkeley CA, USA}
\par {\footnotesize $^{45}$ Laboratório de Instrumentação e Física Experimental de Partículas (LIP), Lisboa, Portugal}
\par {\footnotesize $^{46}$ Laboratoire Leprince-Ringuet, École Polytechnique, CNRS/IN2P3, Université Paris-Saclay, Palaiseau, France}
\par {\footnotesize $^{47}$ Fakultät für Physik, Ludwig-Maximilians-Universität München, München, Germany}
\par {\footnotesize $^{48}$ Laboratoire de Physique Corpusculaire, Clermont Université and Université Blaise Pascal and CNRS/IN2P3, Clermont-Ferrand, France}
\par {\footnotesize $^{49}$ LPNHE, Université Pierre et Marie Curie, Université Paris Diderot, CNRS/IN2P3, Paris, France}
\par {\footnotesize $^{50}$ Laboratoire de Physique Subatomique et de Cosmologie, Université Joseph Fourier and CNRS/IN2P3 and Institut National Polytechnique de Grenoble, Grenoble, France}
\par {\footnotesize $^{51}$ Department of Physics, University of Massachusetts, Amherst, MA, USA}
\par {\footnotesize $^{52}$ Max-Planck-Institut für Physik (Werner-Heisenberg-Institut), München, Germany}
\par {\footnotesize $^{53}$ Niels Bohr Institute, University of Copenhagen, Kobenhavn, Denmark}
\par {\footnotesize $^{54}$ Nikhef National Institute for Subatomic Physics and University of Amsterdam, Amsterdam, Netherlands}
\par {\footnotesize $^{55}$ Department of Physics, New York University, New York, NY, USA}
\par {\footnotesize $^{56}$ Center for High Energy Physics, University of Oregon, Eugene OR, USA}
\par {\footnotesize $^{57}$ Port d’Informació Científica (PIC), Universitat Autònoma de Barcelona (UAB), Barcelona, Spain}
\par {\footnotesize $^{58}$ High Energy Physics (IHEP), Protvino, Russia}
\par {\footnotesize $^{59}$ STFC Rutherford Appleton Laboratory, Didcot, United Kingdom}
\par {\footnotesize $^{60}$ Department of Physics, Royal Holloway University of London, Surrey, United Kingdom}
\par {\footnotesize $^{61}$ SLAC National Accelerator Laboratory, Menlo Park CA, USA}
\par {\footnotesize $^{62}$ Simons Foundation, New York, USA}
\par {\footnotesize $^{63}$ SUBATECH, IMT Atlantique, Université de Nantes, CNRS-IN2P3, Nantes, France}
\par {\footnotesize $^{64}$ National Taiwan University, Taipei, Taiwan}
\par {\footnotesize $^{65}$ Tata Institute of Fundamental Research, Mumbai, India}
\par {\footnotesize $^{66}$ Technische Universit\"at Dortmund, Dortmund, Germany}
\par {\footnotesize $^{67}$ Department of Physics, University of Alberta, Edmonton, AB, Canada}
\par {\footnotesize $^{68}$ Department of Physics, University of Massachusetts, Amherst, MA, USA}
\par {\footnotesize $^{69}$ Department of Physics, University of Arizona, Tucson, AZ, USA}
\par {\footnotesize $^{70}$ H.H. Wills Physics Laboratory, University of Bristol, Bristol, United Kingdom}
\par {\footnotesize $^{71}$ Department of Physics, University of British Columbia, Vancouver, BC, Canada}
\par {\footnotesize $^{72}$ Department of Physics and Astronomy, University of California Irvine, Irvine, CA, USA}
\par {\footnotesize $^{73}$ University of California, San Diego, La Jolla, USA}
\par {\footnotesize $^{74}$ Cavendish Laboratory, University of Cambridge, Cambridge, United Kingdom}
\par {\footnotesize $^{75}$ Physics Department, University of Cape Town, Cape Town, South Africa}
\par {\footnotesize $^{76}$ Carleton University, Ottawa, ON, Canada}
\par {\footnotesize $^{77}$ University of Illinois Urbana-Champaign, Champaign, Illinois, USA}
\par {\footnotesize $^{78}$ Enrico Fermi Institute, University of Chicago, Chicago, IL, USA}
\par {\footnotesize $^{79}$ University of Cincinnati, Cincinnati, OH, USA}
\par {\footnotesize $^{80}$ Department of Physics, Duke University, Durham, NC, USA}
\par {\footnotesize $^{81}$ SUPA - School of Physics and Astronomy, University of Edinburgh, Edinburgh, United Kingdom}
\par {\footnotesize $^{82}$ Gangneung-Wonju National University, South Korea}
\par {\footnotesize $^{83}$ Section de Physique, Université de Genève, Geneva, Switzerland}
\par {\footnotesize $^{84}$ SUPA - School of Physics and Astronomy, University of Glasgow, Glasgow, United Kingdom}
\par {\footnotesize $^{85}$ Physikalisches Institut, Ruprecht-Karls-Universitat Heidelberg, Heidelberg, Germany}
\par {\footnotesize $^{86}$ Department of Physics and Astronomy, University of Kentucky, Lexington, USA}
\par {\footnotesize $^{87}$ Physics Department, Lancaster University, Lancaster, United Kingdom}
\par {\footnotesize $^{88}$ Department of Physics, Jožef Stefan Institute and University of Ljubljana, Ljubljana, Slovenia}
\par {\footnotesize $^{89}$ Fysiska institutionen, Lunds Universitet, Lund, Sweden}
\par {\footnotesize $^{90}$ School of Physics and Astronomy, University of Manchester, Manchester, United Kingdom}
\par {\footnotesize $^{91}$ University of Maribor, Ljubljana, Slovenia}
\par {\footnotesize $^{92}$ Department of Physics, The University of Michigan, Ann Arbor, MI, USA}
\par {\footnotesize $^{93}$ Departamento de Física, Universidade do Minho, Braga, Portugal}
\par {\footnotesize $^{94}$ University of Nebraska-Lincoln, Lincoln, USA}
\par {\footnotesize $^{95}$ Department of Physics and Astronomy, University of New Mexico, Albuquerque NM, USA}
\par {\footnotesize $^{96}$ University of Notre Dame, Notre Dame, USA}
\par {\footnotesize $^{97}$ Homer L. Dodge Department of Physics and Astronomy, University of Oklahoma, Norman OK, USA}
\par {\footnotesize $^{98}$ The Ohio State University, Columbus, USA}
\par {\footnotesize $^{99}$ Department of Physics, University of Oslo, Oslo, Norway}
\par {\footnotesize $^{100}$ Universidad de Oviedo, Oviedo, Spain}
\par {\footnotesize $^{101}$ Department of Physics, Oxford University, Oxford, UK}
\par {\footnotesize $^{102}$ Department of Physics and Astronomy, University of Pittsburgh, Pittsburgh, PA, USA}
\par {\footnotesize $^{103}$ Princeton University, Princeton, USA}
\par {\footnotesize $^{104}$ Rice University, Houston, TX, USA}
\par {\footnotesize $^{105}$ Universidade Estadual Paulista, São Paulo, Brazil}
\par {\footnotesize $^{106}$ Dipartimento di Fisica, Università La Sapienza, Roma, Italy}
\par {\footnotesize $^{107}$ School of Physics, Shandong University, Shandong, China}
\par {\footnotesize $^{108}$ Department of Physics and Astronomy, University of Sheffield, Sheffield, United Kingdom}
\par {\footnotesize $^{109}$ University of Sofia, Sofia, Bulgaria}
\par {\footnotesize $^{110}$ Department of Physics, The University of Texas at Arlington, Arlington, TX, USA}
\par {\footnotesize $^{111}$ National Research Tomsk Polytechnic University, Tomsk, Russia}
\par {\footnotesize $^{112}$ Department of Physics, University of Toronto, Toronto, ON, Canada}
\par {\footnotesize $^{113}$ Dipartimento di Chimica, Fisica e Ambiente, Università di Udine, Udine, Italy}
\par {\footnotesize $^{114}$ Department of Physics, University of Warwick, Coventry, United Kingdom}
\par {\footnotesize $^{115}$ University of Wisconsin - Madison, Madison, WI, USA}
\par {\footnotesize $^{116}$ Department of Physics, Yale University, New Haven, CT, USA}
\par {\footnotesize $^{117}$ IOP and GUST, Vietnam Academy of Science and Technology (VAST), Hanoi, Vietnam}
\par {\footnotesize $^{118}$ Virginia Tech, Blacksburg, Virginia, USA}
\bigskip
\par {\footnotesize $^{a}$ Community White Paper Editorial Board Member}
\par {\footnotesize $^{b}$ Vladimir V. Gligorov acknowledges funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme under grant agreement No 724777 “RECEPT”}
\par {\footnotesize $^{c}$ Supported by the US-DOE, DE-AC02-07CH11359}
\par {\footnotesize $^{d}$ Supported by FCT-Portugal, IF/00050/2013/CP1172/CT0002}
\par {\footnotesize $^{e}$ Supported by the US-NSF, ACI-1450319}
\par {\footnotesize $^{f}$ Supported by the US-NSF, PHY-1607578}
\par {\footnotesize $^{g}$ Supported by ES-MINECO, FPA2016-80994-c2-1-R \& MDM-2015-0509}
\par {\footnotesize $^{h}$ Caterina Doglioni acknowledges funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme under grant agreement No 679305 “DARKJETS”}
\par {\footnotesize $^{i}$ Supported by the US-DOE, DE-SC0009973}
\par {\footnotesize $^{j}$ Supported by the ES-MINECO FPA2016-80994-C2-2-R}
\par {\footnotesize $^{k}$ Supported by the US-DOE, DE-SC0018098 and US-NSF ACI-1558233}
\par {\footnotesize $^{l}$ Supported by the US-DOE, DE-SC0007859 and US-NSF, 76749/1136652/2}
\par {\footnotesize $^{m}$ Supported by funding from the European Union’s Horizon 2020 research and innovation programme under the Marie Skłodowska-Curie grant agreement No. 752730}
\par {\footnotesize $^{n}$ Supported by Swiss National Science Foundation, Early Postdoc Mobility Fellowship, project number P2ELP2\_168556}
\par {\footnotesize $^{o}$ Supported by CONACYT (Mexico)}
\par {\footnotesize $^{p}$ Supported by the US-DOE, DE-AC02-76SF0051}