### 2018-03-23 CWP/papers/HSF-CWP-2017-04_doma/latex/authors.tex
CWP authors: add M. Livny
- Also fix affiliation in DOMA paper

\author[2]{Berzano, Dario}
\author[16]{Bianchi, Riccardo Maria}
\author[2]{Bird, Ian}
\author[13]{Bockelman, Brian}
\author[2]{Campana, Simone}
\author[18]{De, Kaushik}
\author[2]{Duellmann, Dirk}
\author[17]{Elmer, Peter}
\author[11]{Gardner, Robert}
\author[15]{Garonne, Vincent}
\author[6]{Grandi, Claudio}
\author[5]{Gutsche, Oliver}
\author[9]{Hanushevsky, Andrew}
\author[5]{Holzman, Burt}
\author[5]{Jayatilaka, Bodhitha}
\author[10]{Jimenez, Ivo}
\author[8]{Jouvin, Michel}
\author[2]{Keeble, Oliver}
\author[1]{Klimentov, Alexei}
\author[4]{Kuznetsov, Valentin}
\author[1]{Lancon, Eric}
\author[2]{Lassnig, Mario}
\author[3]{Livny, Miron}
\author[10]{Maltzahn, Carlos}
\author[12]{McKee, Shawn}
\author[7]{Menasce, Dario}
\author[5]{Norman, Andrew}
\author[17]{Pivarski, Jim}
\author[2]{Schulz, Markus}
\author[14]{Severini, Horst}
\author[10]{Sevilla, Michael}
\author[5]{Sexton-Kennedy, Elizabeth}
\author[5]{Vaandering, Eric}
\author[10]{Watkins, Noah}
\author[1]{Wenaus, Torre}
\author[10]{Wuerthwein, Frank}
\affiliation[1]{Physics Department, Brookhaven National Laboratory, Upton, NY, USA}
\affiliation[2]{CERN, Geneva, Switzerland}
\affiliation[3]{Center for High Throughput  Computing, Computer Sciences Department, University of Wisconsin-Madison, Madison, WI, USA}
\affiliation[4]{Cornell University, Ithaca, NY, USA}
\affiliation[5]{Fermi National Accelerator Laboratory, Batavia, IL, USA}
\affiliation[6]{INFN Sezione di Bologna, Università di Bologna, Bologna, Italy}
\affiliation[7]{INFN Sezione di Milano-Bicocca, Milano, Italy}
\affiliation[8]{LAL, Université Paris-Sud and CNRS/IN2P3, Orsay, France}
\affiliation[9]{SLAC National Accelerator Laboratory, Menlo Park, CA, USA}
\affiliation[10]{University of California, San Diego, La Jolla, CA, USA}
\affiliation[11]{Enrico Fermi Institute, University of Chicago, Chicago, IL, USA}
\affiliation[12]{Department of Physics, The University of Michigan, Ann Arbor, MI, USA}
\affiliation[13]{University of Nebraska-Lincoln, Lincoln, NE, USA}
\affiliation[14]{Homer L. Dodge Department of Physics and Astronomy, University of Oklahoma, Norman, OK, USA}
\affiliation[15]{Department of Physics, University of Oslo, Oslo, Norway}
\affiliation[16]{Department of Physics and Astronomy, University of Pittsburgh, Pittsburgh, PA, USA}
\affiliation[17]{Princeton University, Princeton, NJ, USA}
\affiliation[18]{Department of Physics, The University of Texas at Arlington, Arlington, TX, USA}