### 2018-04-16 CWP/papers/HSF-CWP-2017-04_doma/latex/authors.tex
CWP DOMA: fix author list
- Add a specific footnote for DOMA editors
- Remove mention of CWP EB members

\author[2]{Dario Berzano,}
\author[16]{Riccardo Maria Bianchi,}
\author[2]{Ian Bird,}
\author[13]{Brian Bockelman,}
\author[2]{Simone Campana,}
\author[18]{Kaushik De,}
\author[2,d]{Dirk Duellmann,}
\author[17]{Peter Elmer,}
\author[11]{Robert Gardner,}
\author[15]{Vincent Garonne,}
\author[6]{Claudio Grandi,}
\author[5,a]{Oliver Gutsche,}
\author[9,c,d]{Andrew Hanushevsky,}
\author[5,a]{Burt Holzman,}
\author[5,a,d]{Bodhitha Jayatilaka,}
\author[10]{Ivo Jimenez,}
\author[8,d]{Michel Jouvin,}
\author[2,d]{Oliver Keeble,}
\author[1]{Alexei Klimentov,}
\author[4]{Valentin Kuznetsov,}
\author[1]{Eric Lancon,}
\author[2]{Mario Lassnig,}
\author[3]{Miron Livny,}
\author[10]{Carlos Maltzahn,}
\author[12,b]{Shawn McKee,}
\author[7]{Dario Menasce,}
\author[5,a]{Andrew Norman,}
\author[17]{Jim Pivarski,}
\author[11]{Benedikt Riedel,}
\author[2]{Markus Schulz,}
\author[14]{Horst Severini,}
\author[10]{Michael Sevilla,}
\author[5,a]{Elizabeth Sexton-Kennedy,}
\author[5,a]{Eric Vaandering,}
\author[11]{Ilija Vukotic,}
\author[10]{Noah Watkins,}
\author[1]{Torre Wenaus}
\author[10]{and Frank Wuerthwein}
\affiliation[1]{Physics Department, Brookhaven National Laboratory, Upton, NY, USA}
\affiliation[2]{CERN, Geneva, Switzerland}
\affiliation[3]{Center for High Throughput  Computing, Computer Sciences Department, University of Wisconsin-Madison, Madison, WI, USA}
\affiliation[4]{Cornell University, Ithaca, NY, USA}
\affiliation[5]{Fermi National Accelerator Laboratory, Batavia, IL, USA}
\affiliation[6]{INFN Sezione di Bologna, Università di Bologna, Bologna, Italy}
\affiliation[7]{INFN Sezione di Milano-Bicocca, Milano, Italy}
\affiliation[8]{LAL, Université Paris-Sud and CNRS/IN2P3, Orsay, France}
\affiliation[9]{SLAC National Accelerator Laboratory, Menlo Park, CA, USA}
\affiliation[10]{University of California, San Diego, La Jolla, CA, USA}
\affiliation[11]{Enrico Fermi Institute, University of Chicago, Chicago, IL, USA}
\affiliation[12]{Department of Physics, The University of Michigan, Ann Arbor, MI, USA}
\affiliation[13]{University of Nebraska-Lincoln, Lincoln, NE, USA}
\affiliation[14]{Homer L. Dodge Department of Physics and Astronomy, University of Oklahoma, Norman, OK, USA}
\affiliation[15]{Department of Physics, University of Oslo, Oslo, Norway}
\affiliation[16]{Department of Physics and Astronomy, University of Pittsburgh, Pittsburgh, PA, USA}
\affiliation[17]{Princeton University, Princeton, NJ, USA}
\affiliation[18]{Department of Physics, The University of Texas at Arlington, Arlington, TX, USA}
\note[a]{Supported by the US-DOE, DE-AC02-07CH11359}
\note[b]{Supported by the US-DOE, DE-SC0007859 and US-NSF, 76749/1136652/2}
\note[c]{Supported by the US-DOE, DE-AC02-76SF0051}
\note[d]{CWP Data Organisation, Management and Access paper Editorial Board member}