### 2018-05-22 CWP/papers/HSF-CWP-2017-01_roadmap/latex/cwp-abstract.tex
Break apart LaTeX for arXiv style
Separate the main text body (cwp-main.tex) from the LaTeX for
the specific document class used for sumbission to arXiv in
JHEP style

\abstract{Particle physics has an ambitious and broad experimental 
programme for the coming decades. This programme requires large 
investments in detector hardware, either to build new facilities 
and experiments, or to upgrade existing ones.
Similarly, it requires commensurate investment in the R\&D of software to acquire, manage, process, and analyse the shear amounts of data to be recorded.
In planning for the HL-LHC in particular, it is critical that all of the
collaborating stakeholders agree on the software goals and priorities,
and that the efforts complement each other. In this spirit, this white paper
describes the R\&D activities required to prepare for this software upgrade.}