### 2018-06-11 CWP/papers/HSF-CWP-2017-02_training/latex/preamble.tex
Training CWP Paper - Import from sharelatex
This is a verbatim import of the document sources from
sharedlatex. Note this version does not compile properly,
which will be fixed in the next commits.

\usepackage[utf8]{inputenc}
\usepackage[USenglish]{babel}
\usepackage{enumitem}
\usepackage{tikz}
\usepackage{verbatim}

\begin{tabular*}{\linewidth}{lc@{\extracolsep{\fill}}r@{\extracolsep{0pt}}}
 & & HSF-CWP-2017-06 \\
 & & June 6, 2018 \\ % use \date or hardwire e.g. December 15, 2017
 & & \\
\end{tabular*}
\vspace{2.0cm}

\title{HEP Software Foundation Community White Paper Working Group --
Training, Staffing and Careers}

%\author{See page~\pageref{sec:author-list} for the full author list}
\date{March 2018}

\input{authors.tex}

\abstract{The rapidly increasing evolution of technology and the parallel increasing complexity of algorithmic analysis in HEP (and in science in general) require a much larger portfolio of specific skills by the developers with respect to the past. Young researchers, graduating from universities worldwide, currently do not receive adequate preparation in the very diverse fields of modern computing to respond to growing necessities posed by the most advanced experimental challenges. The vast majority of the HEP community strongly agrees on the need for an overarching umbrella of training programs to bring young researchers (but also more mature ones) up to date with new software technologies, such as, for example, concurrent programming and artificial intelligence. }

\usepackage{graphicx}
%\usepackage[a4paper,margin=2cm]{geometry}% A4 for a US project?
\usepackage{lineno}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{hyperref}
\hypersetup{
  colorlinks=true,        % false: boxed links; true: colored links
  linkcolor=black,        % color of internal links
  citecolor=green,        % color of links to bibliography
  filecolor=magenta,      % color of file links
  urlcolor=blue           % color of external links
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{amssymb}               % http://ctan.org/pkg/amssymb
\usepackage{pifont}                % http://ctan.org/pkg/pifont
\providecommand{\xmark}{\ding{55}} % https://tex.stackexchange.com/questions/42619/x-mark-to-match-checkmark

%N.B.: If you are going to be defining your own commands please use \providecommand instead of \newcommand. c.f.: https://tex.stackexchange.com/questions/36175/what-do-newcommand-renewcommand-and-providecommand-do-and-how-do-they-differ - Matthew

% https://tex.stackexchange.com/questions/25701/bibtex-vs-biber-and-biblatex-vs-natbib
\usepackage[backend=biber,defernumbers=true,sorting=none,firstinits=true,natbib=true,style=numeric]{biblatex}
\addbibresource{bib/references.bib}