### 2018-06-11 CWP/papers/HSF-CWP-2017-02_training/latex/src/training.tex

% ---------------------------------------------------
% CHAPTER SUMMARY MESSAGE: Motivate the need for training in the HEP community,
%  summarize existing training efforts to motivate that we already spend funding on
% this. Discuss initiatives and programs which can be developed in the coming years
% including a discussion of the known training needs of the community. 
%
%    Goals for this chapter 
%   - Motivation to dedicate resources to training
%   - The community does not recognize the importance
%   - people learn on the job
%   - are there objective criteria
%   - retaining people with skills? 
%
%   see e.g. https://github.com/particleist/inventingcarrots/blob/master/inventing_carrots_merged.md which tries to address the problem of "how do we incentivize affirm and reward learning about software in HEP"
%
%   - People should have incentives to 
%
% ---------------------------------------------------

%Contributors: Fernanda Psihas, Justin Vasel, Michela Paganini, Douglas Davis, Savannah
%Thais, Martin Vala, Jonas Eschle, Xavier Vilasís-Cardona, Jamal Rorie, Sean-Jiun Wang,
%Liz Sexton-Kennedy, Sergei Gleyzer, Laurent Basara, Kurt Rinnert, Bob Stienen, Matt Bellis, Meghan Kane (Industry - SoundCloud), Dario Menasce, Pat Skubic, Hadrien Grasland, Riccardo Maria Bianchi, Daniel S. Katz, Sudhir Malik, Albert Puig

%____________________________________________________________________
%____________________________________________________________________
\section{Introduction}

Throughout the last few decades, the field of high energy physics (HEP) has consistently either driven or utilized the latest innovations in computational tools and technologies for problem solving. 
The toolset of the particle physicist is vast and ever-growing, and the problem set increasingly complex.
A typical graduate student or postdoc working on HEP experiments will likely encounter a number of tools, software, and programming languages with which they are totally unfamiliar. 
This requires ``on-the-job'' training for the same people whose work drives the reach of our analyses, while they are developing the tools for these analyses.

A strong research community should consist of independent actors who ask research questions and direct their own analyses, instead of blindly following prescriptions and recipes. It is therefore imperative that the community be proactive about providing training resources and opportunities for the good of the community, and that we follow through on the promise that adequate career path possibilities exist for people in HEP who would otherwise leave the field due to limited advancement opportunities.

To produce the best possible science, it is important, for the broad HEP community, to invest in training physicists in essential software and software engineering skills, which exist at many different levels. The level of depth of software engineering competences in which a physicist should be trained will vary depending on the problems that they face, and the appropriate level may evolve over time according to the problem at hand. For example, for a researcher who only performs basic data analysis, it may be sufficient to be familiar with just Python and data analysis libraries. However, contributing to a large open source project, such as ROOT, requires greater training in software development and related practices. These higher-order skills may include an understanding of writing testable code, time and space complexity implications, and knowledge of common software architecture paradigms. The need for these skills cannot be fulfilled completely just by hiring specialists in computer science, since a thorough understanding of the physics problems is usually very important to the success of the effort, making it necessary to provide a career path for those young physicists willing to steer their interests in that particular direction.

In each HSF workshop, a discussion on the need for proper tutoring, training, and the relevant career possibilities in HEP computing emerged and was felt to be crucial and essential by almost all participants. There is a strong consensus that future software developments will benefit enormously from a successful implementation of such training programs, provided proper career prospects are put in place, as an incentive, to would-be trainers.

This document tries to address the most relevant questions related to the training in the HEP community, namely:
\begin{itemize}
    \item How to encourage and provide incentives to the members of the HEP community to train students and/or other collaborators?
    \item How to properly assign credit to software development (and training thereof) activities?
    \item What are the current practices in the field regarding hiring of researchers with a strong specialisation in computing and the policies to retain such an expertise in the long term? Is there room for improvement?
    \item How to address and bridge the gap in formal software training that is not always imparted by universities?
\end{itemize}

%____________________________________________________________________
%____________________________________________________________________
\section{Motivation}

%    DOES THIS SECTION ANSWER THE FOLLOWING: 
%    - How is training different in software? Specify that there is /no formal training for these tools?

The introduction and development of increasingly complex tools has led to increasing improvements in the performance of our analyses. However, the personnel who drive these advancements through novel implementations of the technology are the analyzers. In this context, the term {\it analyzers} includes all those researchers who participate in an HEP experiment whose contribution includes designing and implementing a high speed DAQ system, setting up a modern database to store the state of a detector, complex task of managing the data flow from the first trigger level to the final n-tuple dataset and creating sophisticated statistical and mathematical tools to extract publishable physical meaning from those same data.

The development of new tools is insufficient to drive advances in the field unless researchers who have the expertise apply them to physics questions. Furthermore, a larger and a better-connected community, with expertise in the tools and access to the new developments, can lead to an open-source model in HEP. The efforts to train the community in all relevant tools is, thus, a priority within the R\&D effort itself.

However, as tools become more complex, the learning rate of developers must keep up if we are to utilize them effectively.  At the same time, as in other branches of research, it is essential to sustain efforts in maintaining reproducibility of the usage of the tools. This again highlights the need for physicts to be appropriately trained in computing science, rather than professional computing scientists, whose goals and interests may not necessarily be aligned or focused on the physics results of a particular HEP experiment.

To avoid duplicate efforts within the community, it is important to find a way to approach training efforts holistically and cross-experimentally, whenever possible. This requires a certain level of organization and coordination in order to guarantee the delivery of efficient, non-redundant software to different experiments and to allow the different communities to agree on common standards and procedures.

Beyond the immediate benefit to the scientific community in having well-trained collaborators, most people who start a physics graduate program will have careers outside of academia. This trend is reflected in several calls for proposals from national scientific funding agencies like H2020 and those from the National Science Foundation (NSF) that promote training in computer infrastructure and information areas.

%\textcolor{red}{this call is more to develop cyberinfrastructure staff than scientists - for example, people at computing centers}

\begin{quote}
\it The overarching goal of these programs is to prepare, nurture and grow the national scientific workforce for creating, utilizing, and supporting advanced cyberinfrastructure (CI) that enables cutting-edge science and engineering and contributes to the Nation's overall economic competiveness and security.

{\bf - Training-based Workforce Development for Advanced Cyberinfrastructure (CyberTraining) Webinar}\footnote{ \url{https://www.nsf.gov/events/event_summ.jsp?cntn_id=190179&org=NSF}}
\end{quote}

\begin{quote}
\it New information, communication, and computational technologies have had profound impacts on the practice of science (in this solicitation, the term science includes the natural, mathematical, computing, and social sciences),  engineering, and education. This includes the means by which citizens of all ages use science  and engineering to enhance professional and private lives.  The systems, tools, and services emerging from these new technologies are linked to create a comprehensive cyberinfrastructure that is enabling individuals, groups, and organizations to advance research and education in ways that revolutionize who can participate, what they can do, and how they do it. Sustaining this revolution across all areas of science, engineering, and education requires the formation of a citizenry and workforce with the knowledge and skills needed to design and deploy as well as adopt and apply these cyber-based systems, tools and services over the long-term. The opportunity for such preparation should be available at all stages of formal and informal education (K-16 and lifelong), training and professional development, and must be extended to all individuals and communities.

{\bf - Cyberinfrastructure Training, Education, Advancement, and Mentoring for Our 21st Century Workforce (CI-TEAM)}\footnote{ \url{https://www.nsf.gov/funding/pgm_summ.jsp?pims_id=12782}}
\end{quote}

An open issue is the kind of incentive that our community can collectively leverage in order to increase the number of people (with strong expertise in computing) who decide to dedicate time and effort to train fellow members of their community. 
To be effective, the training must be done by people who work at the cutting edge of the technology, who in many cases are found among the youngest researchers. These are also the very same people who are most in need of officially recognized credits for the advancement of their careers. The current practice in HEP is that these young scientists only receive credit and recognition (in the vast majority of cases) by working on a data analysis project, and this discourages them from finding time to transfer their acquired know-how to others in an effective way. A possible solution to this conundrum could be the establishment of specific career paths for researchers with a computer science specialization. The practical implementation of such an approach, however, is extremely challenging as there are a wide range of institutions involved that belong to different countries with their own policies, priorities, and funding strategies. Broadly, two career paths could be envisaged. Both would recognise that domain-sepcific knowledge and continued engagement is vital to the success of HEP computing and software. While in an ideal world (for physicists, at least), a physicst could make a detailed plan of what they need (expressed in a tidy requirements document) and a computer scientist could use this document to provide the required solution, with very little additional exchange of information during the process (which is how the software development process sometimes works in industry). In the real HEP world, instead, most of the time the requirements are bound to change rather frequently for a variety of reasons (not only because of poor planning) forcing developers and physicsts to closely interact during the entire process of software development; this works smoothly only when both communities have the same goals and speak the same jargon, hence the need for physicists with a computer science specialisation.

The first possible career would be the path of a \emph{physicist researcher with computing science specialisation}, in some places corresponding to the role of a physicist-programmer. It would be conceptually different to that of a \emph{computer scientist}: the two have different goals, seek different paths to the solution of their problems, and usually do not even adopt the same jargon in expressing problems and solutions. Such a role would have an active role in physics analysis, and so meet many of the traditional metrics for an academic career path.

The second would be the path of a domain-specific \emph{researcher software engineer}. Such a person has the primary role of developing software and finding software solutions, but with a large domain-specific knowledge to both understand the use cases and clients, but also of the available solutions.  Such posts are almost impossible to develop in the current academic frameworks; but while RSEs are understood by many funding agencies and universities as a technical post, they are seen as available for short-term engagements with a researcher community, delivering a quick fix or a refactoring and then moving one. This does not deliver the domain specific expertise required.

Both career paths suffer in academic terms from few or poorly cited publications, few opportunities to win grants in their own right and little opportunity for impact with commerce and wider society. These three failings make an academic appointment challenging or disfavoured compared to other applicants. However, there are measures that an be taken to address all three weaknesses. 
%This makes the development of specific solutions even more difficult and cumbersome.

%Finally, the career of these two figures follows slightly different paths, with different requirements. 
%____________________________________________________________________
%\subsection{Existing Training efforts  \textcolor{pink}{is this needed?....}}
% why not, shows what is done already

%____________________________________________________________________
%____________________________________________________________________
\section{Training needs of the community}

% DOES THIS SECTION ANSWER THE FOLLOWING: 
% - Different levels of training are required

The HEP user community consists of a set of users with diverse software experience, interests, and time availablity to learn new techniques.  Any training program must take into account the variation in target audiences. For example, an undergraduate doing a summer research project has different needs and skills than their mentoring professor.
For purposes of training, it is useful to think of all our user community as \emph{students}. More specifically, we can broadly identify four different sets of students, classified by experience level, recognizing that even within these groups there will be people who will learn at different speeds and with differing depths of understanding:

\begin{itemize}[itemsep=0pt,parsep=0pt,topsep=5pt,partopsep=0pt]
   \item \emph{Beginners}: New collaborators with no knowledge of the tools or techniques they are expected to use. These might be undergraduates but they could also be senior researchers who have not yet or not recently engaged in software analysis. These are people in need of some kind of formal training in modern computing techniques (basically compiled and scripting languages together with operating system basics such as filesystems, version control, and command-line shells).
   \item \emph{Intermediate users}: HEP analyzers with some experience in either analysis concepts and/or tools, but  looking to supplement their experiences with more recent and modern approaches. 
   \item \emph{Advanced users}: HEP analysis experts who have mastered current technologies and implementations (at a certain point in time) and want to stay up-to-date with new advanced developments.
   \item \emph{Software developers}: HEP scientists in charge of software development in areas not limited only to analysis, such as DAQ systems, computing infrastructure, databases, pattern recongition, and complete frameworks.
\end{itemize}

Each of these groups has vastly different training needs. However, \emph{whenever possible}, any training program should take advantage of developments in pedagogy, such as active learning\footnote{\url{http://www.crlt.umich.edu/tstrategies/tsal}}, peer learning\footnote{\url{https://en.wikipedia.org/wiki/Peer_learning}}, and web-based training.\footnote{See \S\ref{sec:initiatives}.} In some cases, it may even be advantageous to hand out code samples that are purposely broken or flawed, and ask students to fix or improve them. Learning the material in a way that sticks with the students is difficult and challenging for both the students and the instructor and often takes more time than we would prefer. However, this is the best way to ensure an educated community of scientists who can fully contribute to the physics programs at large, which is really the ultimate goal of any training program.

%____________________________________________________________________
\subsection{Knowledge that needs to be transferred}

At all stages of software and computing training, we should take care to encourage \emph{Good Practices Across the Community} (GPAC), such as error checking, modularity of code design, version control, writing tests, etc. All the key concepts addressed in the training should not be specific to a particular experiment or field of application, but general enough to be useful to the whole HEP community and possibly beyond. In this section, we present a list of specific concepts that need to be taught to members of the community, in order to guarantee the base level of competence needed to write efficient code for the different tasks performed in HEP experiments.

An essential knowledge base that needs to be transferred includes basic programming concepts, data structures, basics of code design, error checking, code management tools, validation and debugging tools. More advanced topics include modularity of code design, advanced data structures, evaluation metrics, writing tests and working with different types of hardware accelerators. Cutting-edge approaches such as the use of hardware accelerators (whether GPUs or FPGAs), Deep Neural Networks, and Machine Learning should be considered important training topics. Additionally, special emphasis should be made on reporting results and documenting them.
The training subjects that the community considers important to pursue are listed as follows:

\begin{itemize}[itemsep=0pt,parsep=0pt,topsep=5pt,partopsep=0pt]
   \item Basic and Advanced Programming Concepts
   \begin{itemize}[itemsep=0pt,parsep=0pt,topsep=5pt,partopsep=0pt]
      \item Object oriented paradigm
      \item Compiled langauges (C++)
      \item Scripting languages (Python, Javascript,...)
   \end{itemize}
   \item Algorithmics 
   \begin{itemize}[itemsep=0pt,parsep=0pt,topsep=5pt,partopsep=0pt]
      \item Boost library
      \item STL algorithms for containers
      \item R and/or ROOT
   \end{itemize}
   \item Existing frameworks (development or application level)
   \begin{itemize}[itemsep=0pt,parsep=0pt,topsep=5pt,partopsep=0pt]
      \item Qt
      \item ROOT
      \item experiment specific framework (possibily if of potential interest outside the native experiment)
   \end{itemize}
   \item Code design (design patterns)
   \item Development tools
   \begin{itemize}[itemsep=0pt,parsep=0pt,topsep=5pt,partopsep=0pt]
      \item IDEs (Integrated Development Environment)
      \item Debuggers
      \item Profilers
   \end{itemize}
   \item Evaluation metrics
   \item “Trust” metrics such as data driven tests
   \item Specific software implementation training
   \item Good practices 
   \item Code style and clarity
   \item Scripting and data cleaning
   \item Reporting results reproducibly
   \item Writing Documentation
\end{itemize}

%____________________________________________________________________
%____________________________________________________________________
\section{Challenges}

Within a single experiment, different skill sets are needed. In addition to a skill set that contains basic programming language knowledge, testing and code management tools, and experiment-specific framework knowledge, there are more specialized skills that only a subset of the community needs to know, such as software optimization, or low-level hardware interfaces.
While there already are experiment-specific training efforts in place, there are many needs that are in common between experiments. Therefore, we should probably strive to extract the common knowledge and build a common training program. This would reduce duplication of efforts and enable growth of a shared training culture by accumulating and sharing expertise across different experiments.

To realise this goal, a possibility could be the institution of a {\em federation} of existing training initiatives, a very lightwheight umbrella just aimed at improving the efficiency and cost-effectivness  
of this important activity.

\section{Implementation Roadmap}
The implementation of training should target multiple training formats and knowledge transfer, such as videos, wikis, lectures, jupyter notebooks and advanced visualizations, etc.
People who learn in non-conventional ways should not be excluded from the opportunity to learn. Since there are different stages of knowledge transfer, programs should be developed that encourage experts to share knowledge. Issues considered in this section include global training initiatives, strategies for leveraging existing forums, and training tool development.

An important point to consider is the difficulty, often experienced in the past, in developing large software programs across different experiment communities. It is always challenging to organize a Working Group for training across experiments or even across diverse communities (such as HEP and Astroparticle, for example). This is a point where an appropriate incentive program to reward those who train the community might help.

%____________________________________________________________________
\subsection{Initiatives for future training programs\label{sec:initiatives}}

Some methods that can be used for global training include Massive Online Open Courses, hands-on workshops, online knowledge bases, expert trainer volunteer networks, and web-based training approaches.

Massive Online Open Courses (MOOCs) can be used to develop an open-source set of tutorials and tools. Existing online courses such as Udacity and Coursera can be evaluated and recommended to the community. A novel approach such as WikiToLearn could also explored to assess potential benefits (as has already been attempted in the context of the GridKa School of Computing, see below).

Experiment-specific and global knowledge bases can be established with incentives for experts to contribute. They can be open-source so that a lot of knowledge can be added by the trainees themselves as information is learned (this is the particular context where an approach such as WikiToLearn could be of great help).

Question-and-answer websites such as Stack Overflow for HEP are also a very useful resource for common problems and questions. This approach has already been considered by HSF start-up members, and turned out to be difficult to pursue for a number of reasons, but in the future, boundary conditions might change, making this approach viable.

Hands-on workshops are an invaluable part of learning how to apply theoretical concepts in practice. Identifying which of the current workshops are productive and useful, and if they cover all the topics that need to be transferred and that are in demand by the students, is an important action item.

Creating an expert tutor volunteer network is another way to provide training and support to the community. 
This of course requires proper recognition, at least in terms of career prospects as an incentive for young researchers. It is clear that the best possible tutors are, in principle, those people who are actively engaged in modern software developments: these are in general young researchers in the first steps of their careers and, in order to be attracted to devoting substantial time to training and tutoring, they must be assured that such an effort would be properly recognised in an official way. Such kind of recognition is not currently implemented, at least not in a standardized or official way. A possible structure for such a network could be the establishment of a {\em federation} of existing schools.

It should be considered that some professors, who act as graduate student advisors, might need to be encouraged to make sure their students are properly trained. Sometimes, students are instead pushed to learn the bare minimum to get the work done, at the expense of a broader training/education curriculum that could actually yield improved results, but further down the line. One incentive would be to provide training programs that also count as course credit, perhaps as an elective. This model is already in limited use with some online solutions\footnote{\url{https://www.edx.org/credit}}, but this is not universal. Discussions should be started with collaborators at higher education institutions to see what the roadblocks or opportunities would be for these training sessions to serve double duty. 
It should also be considered that not all students will end their career in research or academia: their contribution to the research activity, as students, should therefore also provide them knowledge, know-how, and skills considered a valuable asset by industry, in order to increase their chances of a career outside research.

Difficulties that have emerged in the past with respect to implementing training courses are the lack of funding and the lack of available time by experts in the field. People with enough expertise or insight in the field usually don't have time to devote to prolonged periods of student training, and, even when they can find time, the cost of setting up a training course in an effective way is often beyond what's made available by funding agencies (funds for travel, hosting, setting up a room with a computing infrastructure to allow interactive hands-on session, etc.) A possible solution is a completely different approach to training (but complementary to the already existing and successful classical efforts such as the CERN School of Computing's Bertinoro and Kit ones): instead of directly teaching to students, trainees could make use of a web-based platform to provide training materials to students. This complementary approach has several advantages over traditional ones:
\begin{itemize}[itemsep=0pt,parsep=0pt,topsep=5pt,partopsep=0pt]
   \item Tutors can add material to the web site at a very slow pace (whenever they find time to do it, one slide a week or a chapter a day)
   \item Their material, made public on a web-site, can be further expanded by collaborators (also at their own pace) or, even better and more productively, by students who decide to contribute new additions, examples, exercises etc. Such a collaborative effort allows more people to be exposed to training at any given time, creates a sense of community, and creates bridges between people in contiguous areas or research. Students can use the same platform to exchange their own examples, make suggestions and point out interesting concepts. In such a model, the possible contribution from others to the training material needs to be moderated and validated with appropriate policies.
   \item If complemented by the availability of remote virtual machines (possibly through a browser), students could have access to examples and exercises that are already embedded in their own natural environment: all the necessary tools/libraries needed to implement the exercise will be already available in the virtual machine (possibly a Docker container). With just a web browser, students could run complex examples from home, taking advantage of a remote facility that provides some storage and computing power. Important here is the concept of ``environment'': a Docker container could be set up in such a way that students will work in an exact replica of the environment they will be exposed to in their experiment (or collaboration at large). Moreover, students could be provided with Docker containers that preserves their modified environment across sessions, allowing them to develop their skills over a prolonged period of time by accessing all the files that were made persistent day by day during the training.
   \item There would no longer be a need to find the resources to host a school and pay the tutor(s) (and eventually subsidize the students to participate in training in a remote location). Students could follow the training at their own pace from wherever they happen to be. A traditional school only lasts for 5 days (10 at most) and it is difficult to cover a  subject to any significant depth in such a short time. The web approach, instead, would allow for very long and in-depth coverage of any kind of subject, and in this sense it could be a {\it complementary} approach to a traditional school. Of particular interest could be courses such as, e.g., ``Machine Learning'', ``Statistical Analyis with ROOT'', or even just ``Good practices in C++'' or ``Python Programming for scientific computing''. 
   \item Finally, this approach could allow the creation of {\it browsable} repositories of all training materials, grouped by argument, by relevance, by experiment or whatever other critieria. Students from all over the world could be exposed to a large repository of examples, exercises and in general training material from their own home.
\end{itemize}

An example of a such a web-based platform already exists, and has been implemented as an Open Source project (backed by Wikimedia) by a group of more than 30 Italian students. The project is WikiToLearn \footnote{\url{https://it.wikitolearn.org/}}. It hosts training material in several languages, for several disciplines, ranging from Economics to Physics, Mathematics, and several others. Because it is based upon wikimedia\footnote{\url{https://www.wikimedia.org/}} software, it is very easy to add material to the site, and to make it appear under a specific topic (such as, e.g., Software/Techniques/Machine-Learning) and to manipulate it as if it were a single document. In the end, students can selectively choose individual chapters from the site and have the corresponding pdf sent them as a book, complete with index, content, and chapters.

The adoption of such an approach is made rather easy in WikiToLearn by the relative simplicity of the wikimedia-based toolset: users contribute their training material using just a web-browser, and in order to do this efficiently, the learning curve has been kept appropriately shallow.

An interesting exercise in this context has recently been made by colleagues of the GridKa School of computing: the material from this year (2017) has been made publicly available on WTL\footnote{\url{https://en.wikitolearn.org/GridKa2017}}. This constitutes an intersting example of what can be accomplished using this platform; it is just a first example of what's possible, but an inspiring one.

Finally it should be important to evaluate, if and to which extent, the complementary approaches to training (such as schools and dedicated web-site courses) could be of mutual benefit (how to make them cooperate in the development of a complete training program).

  % \item Question-and-answer web sites such as Stack Overflow 
   %(https://en.wikipedia.org/wiki/Stack_Overflow) 
   %/ Stack Exchange are very useful. 
   %(https://en.wikipedia.org/wiki/Stack_Exchange)
  % \item An expert/tutor volunteer network (office hours style) can be established.
   %\item Investing in periodic in-person, hands-on workshops can be continued.

%____________________________________________________________________
\subsection{Leveraging existing forums}

To achieve our goals for training the community, we can take advantage of existing training forums. Resources such as conferences, workshops, and schools (in person and online) can provide a lot of value for our training purposes with little effort to set up. We should leverage the existing training forums that most closely match the HEP community's needs.

Within the HEP community, there are already some working examples of dedicated training environments that alternate between general topics and experiment-specific topics. The LHC Physics Center (LPC) at Fermilab hosts Hands-on Advanced Tutorial Sessions (HATS)\footnote{\url{http://lpc.fnal.gov/programs/schools-workshops/}} throughout the year to introduce and train participants in topics as diverse as the latest $b$-tagging algorithms, Git/GitHub, and machine learning. These HATS provide face-to-face time with instructors and participants at Fermilab, and also allow remote collaborators to join in by video and complete the same online exercises. A similar approach is in use in the CMS Data Analysis Schools (CMSDAS)\footnote{\url{http://lpc.fnal.gov/programs/schools-workshops/cmsdas.shtml}}, a series of week-long workshops that now take place at multiple labs all over the world and are designed to ramp up new collaborators in CMS-specific analysis tools while providing some discussion of the physics as well.

Other examples are CERN School of Computing\footnote{\url{https://csc.web.cern.ch/}}, CERN OpenLab Software workshops\footnote{\url{http://openlab.cern/}} education in collaboration with industry partners, and a series of more advanced topical training courses provided by MPI Munich and DESY that focus on advanced programming, use of acceleration hardware and statistical tools including machine learning. This list includes the Bertinoro\footnote{\url{https://web.infn.it/esc17/index.php?lang=en}},  GridKa\footnote{\url{http://gridka-school.scc.kit.edu/2017/}}, and CODAS-HEP\footnote{\url{http://codas-hep.org}} Schools of Computing.

%Anecdotally, these are viewed as a great success by the CMS community in preparing new collaborators and might serve as a model for future efforts in ML training.

Over the past decade, massive open online courses (MOOCs) have been developed by universities and private organizations. They have been well received by industry and academia.% and are maintained by the outside community.
In addition, they provide a lot of flexibility in terms of cost and time constraints; they are typically free and open for enrollment at any time of the year. Since the material can be accessed at any time and revisited at any time, they can be completed at a pace that makes sense, for example, for a physicist who needs to learn machine leanrning in a piecemeal way.

There are a growing set of MOOCs teaching various subjects. Since there are many options, there is a wide variety with respect to the depth of the material and specific tools taught. Exploring these options allows us to choose which is the right offering for the knowledge needed to work on a specific HEP experiment. We can pick and choose modules to tailor an appropriate roadmap of skills to learn.
More difficult will be to assemble HEP specific training material not already avaible elsewhere in an efficient and organized way, since this requires adequate organization, volounteers and and a suitable infrastructure.

Several industry conferences already exist that bring together those in academia and industry who are at the cutting edge of these techniques. Conferences such as NIPS\footnote{\url{https://nips.cc}} and PyData\footnote{\url{https://pydata.org}} provide a focused place where attendees can learn a lot about machine learning in a short period of time. Machine learning concepts such as current methods, tools, and problems facing industry and academia can be learned at conferences. In addition, conferences are an excellent networking opportunity; attendees can meet and share ideas with fellow learners and experts. Bonds can be formed quickly at conferences that can be maintained after the duration of the conference. These connections to the outside community can be essential since we will be evolving training materials to ensure that they stay relevant over time.

For example, at the time of this writing, Coursera\footnote{\url{https://www.coursera.org/learn/machine-learning}} and Udacity\footnote{\url{https://www.udacity.com/course/intro-to-machine-learning--ud120}} both provide great machine learning massive open online courses at no cost. These two courses both provide a great foundation for learning machine learning fundamentals. However, Coursera's approach emphasizes more theory (with more math background necessary) and uses MATLAB/Octave while Udacity's approach emphasizes practical using machine learning techniques in a practical sense using Python tooling.

%____________________________________________________________________
\subsection{Resources and Incentives}

Training the community is something that a large cross-section of the community understands to be important, but finding time and effort to contribute to this effort is not actively on the radar of most potential contributors. Providing incentives for their participation and creating the appropriate platforms can go a long way to reach a productive training environment. It can be as simple as inviting someone to give a software tutorial on the subject that they are familiar with, give a lecture or seminar or contribute to a growing knowledge base.

Another important incentive is recognition. For younger members of the community, having the opportunity to create a training resource, such as a software tutorial or a knowledge base on a particular topic, is very empowering and motivational to continue the efforts of training others. Engaging younger members of the community is crucial to long-term success of HEP training endeavors.

It is also critical to incorporate training into grant proposals so that it can be connected with other areas such as research and development. Efforts like DIANA-HEP \footnote{\url{http://diana-hep.org/}} and aMVA4NewPhysics \footnote{\url{https://amva4newphysics.wordpress.com/}} that combine training and software development are good examples of such ideas in practice. More examples of such efforts are needed.

\section{Other resources}

Software Carpentry\footnote{\url{https://software-carpentry.org}} and Data Carpentry\footnote{\url{http://www.datacarpentry.org}} are two parts of The Carpentries organization that collaboratively build and teach some of the basic concepts in developing and maintaining software, and analyzing data, respectively.  The materials that they develop and use are open, and can be customized for science domains
(e.g., HEP), or participant groups (e.g., undergraduates). Their model is that they offer training of trainers, and then the trainers who have graduated can offer training under the SC/DC names, though of course, anyone can use the SC/DC materials without doing it under the SC/DC names.
Subjects covered in a typical Data Carpentry school are given below. This list is taken from the curriculum of the CODATA-RDA Research Data Science Summer School in progress in Trieste, Italy during July 2017. (http://indico.ictp.it/event/7974/):

\begin{itemize}
\item Introduction
\item UNIX Shell programming (Software Carpentry Module)
\item GitHub (Software Carpentry Module)
\item R (Software Carpentry Module)
\item BYOD (Bring Your Own Data) best practices
\item Data Management with SQL  (Software Carpentry Module)
\item RDM Storage Management
\item Visualisation
\item Machine Learning Overview - Recommendation
\item Recommender Systems
\item Artificial Neural Networks and other Machine Learning Systems 
\item Research Computational Infrastructure 
\end{itemize}

While Software Carpentry is leveraged to build a foundation of knowledge for later more advanced concepts, it is important to note much of this material is developed specifically for this course and not a part of a larger Software Carpentry program. This course focuses on a shallow but wide building of foundational skills approach, introducing many base concepts but not going deep into any one concept, with a through line of Open Science and Ethical Data Usage.

\section{Conclusions}

The HEP community by and large acknowledges and recognizes the great importance of training in the field of scientific computing. This activity should encompass several types of {\it students}, from undergraduates, to young researchers, up to senior physicists, all of them in need of an appropriately designed training path in order to be proficent in their scientific endevours.

We have identified a certain number of problems that need to be overcome in setting up an appropriate training program: 
\begin{itemize}
    \item Costs and relative funding
    \item Incentives
    \item Career paths
    \item Overall organization across experiments, countries and corresponding Funding Agencies
\end{itemize}

For each of these points we provide an overview of the current situation and possible proposals to build a roadmap.

%____________________________________________________________________
%____________________________________________________________________
%\begin{itemize}
%
%   \item How to motivate the community to participate in developing programs? Can some of it come from \& leadership?
%   \item How to deal with competitiveness in the field? Can’t force people to be interested in forming a community. How can we incentivize this? Building a %vibrant, welcoming community may help.
%\end{itemize}

\section*{Aknowledgements}
A.P. acknowledges support by SNF under contract 168169.