### 2018-10-04 CWP/papers/HSF-CWP-2017-14_recotrigger/latex/authors.tex

Albrecht, Johannes$^{1}$;
Bloom, Kenneth$^{2}$;
Boccali, Tommaso$^{3}$;
Boveia, Antonio$^{4,c}$;
De Cian, Michel$^{5}$;
Doglioni, Caterina$^{6,b}$;
Dziurda, Agnieszka$^{7}$;
Farbin, Amir$^{8}$;
Fitzpatrick, Conor$^{9}$;
Gaede, Frank$^{10}$;
George, Simon$^{11}$;
Gligorov, Vladimir$^{12,a}$;
Grasland, Hadrien$^{13}$;
Grillo, Lucia$^{14}$;
Hegner, Benedikt$^{7}$;
Kalderon, William$^{6}$;
Kama, Sami$^{15}$;
Koppenburg, Patrick$^{16}$;
Krutelyov, Slava$^{17}$;
Kutschke, Rob$^{18}$;
Lampl, Walter$^{19}$;
Lange, David$^{20}$;
Moyse, Ed$^{21}$;
Norman, Andrew$^{18}$;
Petric, Marko$^{7}$;
Polci, Francesco$^{12}$;
Potamianos, Karolos$^{10}$;
Ratnikov, Fedor$^{22}$;
Raven, Gerhard$^{16}$;
Ritter, Martin$^{23}$;
Rizzi, Andrea$^{3}$;
Rodrigues, Eduardo$^{24}$;
Rousseau, David$^{13}$;
Salzburger, Andy$^{7}$;
Sexton Kennedy, Liz$^{18}$;
Sokoloff, Michael D$^{24}$;
Stewart, Graeme$^{7}$;
Ustyuzhanin, Andrey$^{22}$;
Viren, Brett$^{25}$;
Williams, Mike$^{26}$;
Winklmeier, Frank$^{27}$;
Wuerthwein, Frank$^{17}$
\bigskip 
\par {\footnotesize $^{1}$ Technische Universit\"at Dortmund, Dortmund, Germany}
\par {\footnotesize $^{2}$ University of Nebraska-Lincoln, Lincoln, NE, USA}
\par {\footnotesize $^{3}$ INFN Sezione di Pisa, Università di Pisa, Scuola Normale Superiore di Pisa, Pisa, Italy}
\par {\footnotesize $^{4}$ The Ohio State University, Columbus, OH, USA}
\par {\footnotesize $^{5}$ Physikalisches Institut, Ruprecht-Karls-Universitat Heidelberg, Heidelberg, Germany}
\par {\footnotesize $^{6}$ Fysiska institutionen, Lunds Universitet, Lund, Sweden}
\par {\footnotesize $^{7}$ CERN, Geneva, Switzerland}
\par {\footnotesize $^{8}$ Department of Physics, The University of Texas at Arlington, Arlington, TX, USA}
\par {\footnotesize $^{9}$ Institute of Physics, École Polytechnique F\'ed\'erale de Lausanne (EPFL), Lausanne, Switzerland}
\par {\footnotesize $^{10}$ Deutsches Elektronen-Synchrotron, Hamburg, Germany}
\par {\footnotesize $^{11}$ Department of Physics, Royal Holloway University of London, Surrey, United Kingdom}
\par {\footnotesize $^{12}$ LPNHE, Université Pierre et Marie Curie, Université Paris Diderot, CNRS/IN2P3, Paris, France}
\par {\footnotesize $^{13}$ LAL, Université Paris-Sud and CNRS/IN2P3, Orsay, France}
\par {\footnotesize $^{14}$ School of Physics and Astronomy, University of Manchester, Manchester, United Kingdom}
\par {\footnotesize $^{15}$ Physics Department, Southern Methodist University, Dallas TX, United States of America}
\par {\footnotesize $^{16}$ Nikhef National Institute for Subatomic Physics and Vrije Universiteit Amsterdam, Amsterdam, The Netherlands}
\par {\footnotesize $^{17}$ University of California, San Diego, La Jolla, CA, USA}
\par {\footnotesize $^{18}$ Fermi National Accelerator Laboratory, Batavia, IL, USA}
\par {\footnotesize $^{19}$ Department of Physics, University of Arizona, Tucson, AZ, USA}
\par {\footnotesize $^{20}$ Princeton University, Princeton, NJ, USA}
\par {\footnotesize $^{21}$ Department of Physics, University of Massachusetts, Amherst, MA, USA}
\par {\footnotesize $^{22}$ National Research University Higher School of Economics, Moscow, Russia;  Yandex School of Data Analysis, Moscow, Russia}
\par {\footnotesize $^{23}$ Fakult\"at f\"ur Physik, Ludwig-Maximilians-Universit\"at M\"unchen, M\"unchen, Germany}
\par {\footnotesize $^{24}$ University of Cincinnati, Cincinnati, OH, USA}
\par {\footnotesize $^{25}$ Physics Department, Brookhaven National Laboratory, Upton, NY, USA}
\par {\footnotesize $^{26}$ Laboratory for Nuclear Science, Massachusetts Institute of Technology, Cambridge, MA, USA}
\par {\footnotesize $^{27}$ Center for High Energy Physics, University of Oregon, Eugene, OR, USA}
\bigskip
\par {\footnotesize $^{a}$ Vladimir V. Gligorov acknowledges funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme under grant agreement No 724777 “RECEPT”}
\par {\footnotesize $^{b}$ Caterina Doglioni acknowledges funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme under grant agreement No 679305 “DARKJETS”}
\par {\footnotesize $^{c}$ Antonio Bovela acknowledges funding for US Department of Energy (Grant DE-SC0011726).}