### 2018-10-04 CWP/papers/HSF-CWP-2017-14_recotrigger/latex/recoCWP.tex

\documentclass[11pt,letterpaper,fleqn]{article}
\usepackage{graphicx}
\usepackage{lineno}
\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  urlcolor=blue,
  citecolor=blue
}
\usepackage{subfigure}
\usepackage{cite}
\usepackage{xcolor,colortbl}
\usepackage{wrapfig}
\usepackage{enumitem} %%% uncommenting this line and associated
%% options only saves about 1/4 page
% enumitem options
%\setlist{nosep}
%\setitemize{wide}
%\setlist[itemize]{leftmargin=*}
%\setlist[enumerate]{leftmargin=*}

\selectcolormodel{rgb}
\definecolor{LHCb dark}{rgb}{0.0000,0.3412,0.6549}%From LHCb logo
\definecolor{UC red}{rgb}{0.8196,0.1176,0.2314} % from logo
\definecolor{brickred}{rgb}{0.8, 0.25, 0.33}
\definecolor{Gray}{gray}{0.85}

%KC see http://en.wikibooks.org/wiki/LaTeX/Page_Layout
\textwidth      6.5 in
\textheight     9.0 in
\hoffset = 0 in
\voffset = 0 in
\headheight = 0 in
\headsep = 0 in
\topmargin = 0 in
\oddsidemargin  = 0.0 in
\evensidemargin = 0.0 in

%KC For ~11 lines per 2 inches. smaller is not easily legible 
\linespread{0.95}

%adjust enumeration spacing
% see http://en.wikibooks.org/wiki/LaTeX/List_Structures#Line_spacing
\let\oldenumerate\itemize
\renewcommand{\itemize}{
  \oldenumerate
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{1pt}
  \setlength{\parsep}{1pt}
}

\def\s2i2{$S^2 I^2$}

\begin{document}
%\linenumbers
\thispagestyle{empty}

\input{titlepage.tex}
\newpage
%\tableofcontents
%\newpage
\setcounter{page}{1}

\input{all.tex}

%\newpage 
%\bibliographystyle{unsrt}
%\bibliography{recoCWP}

\begin{thebibliography}{9}

\bibitem{HSF2017}
HEP Software Foundation,
\textit{A Roadmap for HEP Software and Computing R\&D for the 2020s},
HSF-CWP-2017-01, arXiv:1712.06982v2 (\href{https://arxiv.org/abs/1712.06982}{Link}).
%\url{http://hepsoftwarefoundation.org/activities/cwp.html}.
%\href{http://hepsoftwarefoundation.org/activities/cwp.html}{HSF CWP page}.

\bibitem{RECOCWP}
Software trigger and event reconstruction CWP group, To be published as HSF-CWP-2017-16 (\href{https://github.com/HEP-SF/documents/tree/master/CWP/papers}{Link}).

\bibitem{Campana2016}
S. Campana,
\textit{Presentation to the 2016 Aix-les-Bains ECFA HL-LHC workshop},
Oct 3 2016 (\href{https://indico.cern.ch/event/524795/contributions/2236590/attachments/1347419/2032314/ECFA2016.pdf}{Link}).

\bibitem{ATLAS2015} ATLAS Collaboration, \textit{ATLAS Phase-II Upgrade Scoping Document},
CERN-LHCC-2015-020, LHCC-G-166 (2015) (\href{https://cds.cern.ch/record/2055248}{Link}).

\bibitem{CMS2015}
CMS Collaboration,
\textit{Technical Proposal for the Phase-II Upgrade of the CMS Detector}, 
CERN-LHCC-2015-010, LHCC-P-008, CMS-TDR-15-02 (2015) (\href{https://cds.cern.ch/record/2020886}{Link}).

\bibitem{LHCb2014}
LHCb Collaboration, 
\textit{LHCb Trigger and Online Upgrade Technical Design Report},
CERN-LHCC-2014-016 (2014) (\href{https://cds.cern.ch/record/1701361}{Link}).

\bibitem{ALICE2015}
ALICE Collaboration, 
\textit{Technical Design Report for the Upgrade of the Online-Offline Computing System}, 
(\href{https://cds.cern.ch/record/2011297}{Link}).

\bibitem{Bird2014}
I. Bird, et. al., 
\textit{Update of the Computing Models of the WLCG and the LHC Experiments},
CERN-LHCC-2014-014, (2014), (\href{http://cds.cern.ch/record/1695401/files/LCG-TDR-002.pdf}{Link}).

\bibitem{Brun1996}
R. Brun and F. Rademakers,
\textit{ROOT - An Object Oriented Data Analysis Framework}, 
Proceedings AIHENP'96 Workshop, Lausanne, Sep. 1996, Nucl. Inst. \& Meth. in Phys. Res. A 389 (1997) 81-86. See also \href{http://root.cern.ch}{ROOT homepage}.

\bibitem{Aaij2016}
R. Aaij, et. al., 
\textit{Tesla : an application for real-time data analysis in High Energy Physics}, 
Comput. Phys. Commun. 208 (2016) 35-42 (\href{https://cds.cern.ch/record/2147693}{Link}).

\bibitem{ATLAS2017}
ATLAS Collaboration, \textit{Performance of the ATLAS Trigger System in 2015} ,
Eur. Phys. J.C77 5 317 (2017),
\href{https://inspirehep.net/record/1500696}. ATLAS Collaboration,
\textit{Trigger-object Level Analysis with the ATLAS detector at the Large Hadron Collider: summary and perspectives}, 
ATL-DAQ-PUB-2017-003 (2017) (\href{https://cds.cern.ch/record/2295739}{Link}).

\bibitem{CMS2016}
CMS Collaboration,
\textit{Search for narrow resonances in dijet final states at s√= 8 TeV with the novel CMS technique of data scouting},
Phys. Rev. Lett. 117 031802 (2016), (\href{https://cds.cern.ch/record/2149625}{Link}).

\end{thebibliography}

\end{document}