### 2020-04-30 LHCC/2020/2020-01/ctcs.tex
Break sources into an input file per section

% Copyright (C) 2020, the authors, licence CC-BY-4.0.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Document now locked in Overleaf, transition to GitHub in progress %%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% JHEP preprint template
\documentclass[11pt,a4paper]{article}
\usepackage{jheppub}
%\notoc

\usepackage[utf8]{inputenc}

\usepackage{graphicx}
\usepackage{subcaption}
\usepackage[titletoc]{appendix}
\setcounter{tocdepth}{2} % Only to subsection level in ToC
\usepackage{xspace}

% Stop lists from using so much space
\usepackage{enumitem}
\setlist{noitemsep}

% References
\usepackage[style=numeric-comp,sorting=none]{biblatex}
\addbibresource{ctcs.bib}

%% Comment out before final submission %%
\usepackage{lineno}  % for line numbering during review
\linenumbers

\usepackage{comment}

\begin{document}

% HSF Document Number (only for arXiv)
\noindent
\begin{tabular*}{\linewidth}{lc@{\extracolsep{\fill}}r@{\extracolsep{0pt}}}
 & & HSF-DOC-2020-01 \\
 & & Zenodo DOI Here \\
 & & May 1, 2020 \\ % use \date or hardwire e.g. December 15, 2017
 & & \\
\end{tabular*}
\vspace{2.0cm}

\title{HL-LHC Computing Review:\\Common Tools and Community Software}

\author{High-Energy Physics Software Foundation}

\maketitle

% Content is stored in separate TeX files for managability
\input{input_introduction}

\input{input_generators}

\input{input_simulation}

\input{input_recoTrigger}

\input{input_analysis}

\input{input_summary}

\sloppy
\raggedright
\clearpage
\printbibliography[title={References},heading=bibintoc]

\end{document}